---
title: "Comparison by Cellchat - multiple sample"
author: "Zack"
date: "2026-06-14"
output: html_document
---

```{r}
library(SpatialCellChat)
library(CellChat)
library(patchwork)
library(ggplot2)
library(future)
library(ComplexHeatmap)
library(grid)
library(Seurat)

future::plan("sequential")
options(future.globals.maxSize = 4 * 1024^3)  # 4 GB

# progressr::handlers(global=TRUE)
```

# load calculated rds

```{r}
# Contact-dependent objects
chat_control_fvb <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat/cellchat_control_multi_contactRange20_final.rds"
)

chat_disease_fam <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat/cellchat_disease_multi_contactRange20_final.rds"
)

# Secreted-signaling objects
chat_control_fvb_secreted <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat/cellchat_control_multi_secreted_range250_final.rds"
)

chat_disease_fam_secreted <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat/cellchat_disease_multi_secreted_range250_final.rds"
)
```

# Merge

```{r}
object.list.contact <- list(
  Control = chat_control_fvb,
  Disease = chat_disease_fam
)
object.list.secreted <- list(
  Control = chat_control_fvb_secreted,
  Disease = chat_disease_fam_secreted
)
cellchat.contact <- mergeSpatialCellChat(
  object.list.contact,
  add.names = names(object.list.contact),
  cell.prefix = TRUE
)
cellchat.secreted <- mergeSpatialCellChat(
  object.list.secreted,
  add.names = names(object.list.secreted),
  cell.prefix = TRUE
)
```

# Load single-sample RDS for donor-averaged interaction number

```{r}
single_sample_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/single_sample_cellchat"

sample.order <- c("FVB", "FVQ", "FAM", "FCM", "FEP", "FVS")
sample.condition <- c(
  FVB = "Control",
  FVQ = "Control",
  FAM = "Disease",
  FCM = "Disease",
  FEP = "Disease",
  FVS = "Disease"
)

object.list.contact.single <- list(
  FVB = readRDS(file.path(single_sample_dir, "chat_FVB_contactRange20_final.rds")),
  FVQ = readRDS(file.path(single_sample_dir, "chat_FVQ_contactRange20_final.rds")),
  FAM = readRDS(file.path(single_sample_dir, "chat_FAM_contactRange20_final.rds")),
  FCM = readRDS(file.path(single_sample_dir, "chat_FCM_contactRange20_final.rds")),
  FEP = readRDS(file.path(single_sample_dir, "chat_FEP_contactRange20_final.rds")),
  FVS = readRDS(file.path(single_sample_dir, "chat_FVS_contactRange20_final.rds"))
)

object.list.secreted.single <- list(
  FVB = readRDS(file.path(single_sample_dir, "chat_FVB_secreted_range250_final.rds")),
  FVQ = readRDS(file.path(single_sample_dir, "chat_FVQ_secreted_range250_final.rds")),
  FAM = readRDS(file.path(single_sample_dir, "chat_FAM_secreted_range250_final.rds")),
  FCM = readRDS(file.path(single_sample_dir, "chat_FCM_secreted_range250_final.rds")),
  FEP = readRDS(file.path(single_sample_dir, "chat_FEP_secreted_range250_final.rds")),
  FVS = readRDS(file.path(single_sample_dir, "chat_FVS_secreted_range250_final.rds"))
)

object.list.contact.single <- object.list.contact.single[sample.order]
object.list.secreted.single <- object.list.secreted.single[sample.order]

get_count_matrix <- function(chat) {
  cell.types <- c("eEVT", "Endothelial", "EVT")
  mat <- as.matrix(chat@net$count)
  mat.full <- matrix(
    0,
    nrow = length(cell.types),
    ncol = length(cell.types),
    dimnames = list(cell.types, cell.types)
  )
  
  source.use <- intersect(cell.types, rownames(mat))
  target.use <- intersect(cell.types, colnames(mat))
  mat.full[source.use, target.use] <- mat[source.use, target.use, drop = FALSE]
  mat.full
}

average_count_matrix <- function(object.list) {
  out <- lapply(c("Control", "Disease"), function(condition.use) {
    sample.use <- names(object.list)[sample.condition[names(object.list)] == condition.use]
    mats <- lapply(object.list[sample.use], get_count_matrix)
    Reduce("+", mats) / length(mats)
  })
  names(out) <- c("Control", "Disease")
  out
}

plot_donor_average_summary <- function(object.list, title.name, measure = c("number", "strength")) {
  measure <- match.arg(measure)
  
  df <- data.frame(
    sample = names(object.list),
    condition = sample.condition[names(object.list)],
    value = sapply(object.list, function(x) {
      if (measure == "number") {
        sum(get_count_matrix(x))
      } else {
        sum(as.matrix(x@net$weight))
      }
    })
  )
  
  mean.df <- aggregate(
    value ~ condition,
    data = df,
    FUN = mean
  )
  
  y.title <- ifelse(
    measure == "number",
    "Mean interaction number per donor",
    "Mean interaction strength per donor"
  )
  
  ggplot(mean.df, aes(x = condition, y = value, fill = condition)) +
    geom_col(width = 0.6) +
    geom_point(
      data = df,
      aes(x = condition, y = value),
      inherit.aes = FALSE,
      position = position_jitter(width = 0.08),
      size = 2
    ) +
    geom_text(
      aes(label = signif(value, 2)),
      vjust = -0.4,
      size = 4
    ) +
    labs(title = title.name, x = NULL, y = y.title) +
    theme_classic() +
    theme(legend.position = "none")
}

plot_donor_average_number <- function(object.list, title.name) {
  plot_donor_average_summary(object.list, title.name, measure = "number")
}

plot_donor_average_strength <- function(object.list, title.name) {
  plot_donor_average_summary(object.list, title.name, measure = "strength")
}

plot_average_count_heatmap <- function(object.list, title.name) {
  avg.mat <- average_count_matrix(object.list)
  df <- do.call(
    rbind,
    lapply(names(avg.mat), function(condition.use) {
      x <- as.data.frame(as.table(avg.mat[[condition.use]]))
      colnames(x) <- c("source", "target", "value")
      x$condition <- condition.use
      x
    })
  )
  
  df$source <- factor(df$source, levels = rev(rownames(avg.mat[[1]])))
  df$target <- factor(df$target, levels = colnames(avg.mat[[1]]))
  
  ggplot(df, aes(x = target, y = source, fill = value)) +
    geom_tile(color = "white", linewidth = 0.4) +
    geom_text(aes(label = signif(value, 2)), size = 3) +
    facet_wrap(~ condition, nrow = 1) +
    scale_fill_gradient(low = "grey95", high = "#B2182B", name = "Mean number") +
    coord_fixed() +
    labs(title = title.name, x = "Target", y = "Source") +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(hjust = 0.5, face = "bold")
    )
}

plot_average_count_circle <- function(object.list, title.prefix, label.edge = FALSE) {
  avg.mat <- average_count_matrix(object.list)
  edge.max <- max(unlist(avg.mat))
  
  par(mfrow = c(1, 2), xpd = TRUE)
  
  for (condition.use in names(avg.mat)) {
    CellChat::netVisual_circle(
      avg.mat[[condition.use]],
      weight.scale = TRUE,
      label.edge = label.edge,
      edge.weight.max = edge.max,
      edge.width.max = 12,
      title.name = paste0(title.prefix, " mean number - ", condition.use)
    )
  }
}

coarse_count_matrix <- function(mat) {
  group.use <- c(eEVT = "EVT", Endothelial = "Endothelial", EVT = "EVT")
  group.levels <- c("EVT", "Endothelial")
  out <- matrix(0, nrow = 2, ncol = 2, dimnames = list(group.levels, group.levels))
  
  for (source in rownames(mat)) {
    for (target in colnames(mat)) {
      out[group.use[source], group.use[target]] <- out[group.use[source], group.use[target]] + mat[source, target]
    }
  }
  
  out
}

average_coarse_count_matrix <- function(object.list) {
  out <- lapply(c("Control", "Disease"), function(condition.use) {
    sample.use <- names(object.list)[sample.condition[names(object.list)] == condition.use]
    mats <- lapply(object.list[sample.use], function(x) coarse_count_matrix(get_count_matrix(x)))
    Reduce("+", mats) / length(mats)
  })
  names(out) <- c("Control", "Disease")
  out
}

plot_average_coarse_count_circle <- function(object.list, title.prefix) {
  avg.mat <- average_coarse_count_matrix(object.list)
  edge.max <- max(unlist(avg.mat))
  
  par(mfrow = c(1, 2), xpd = TRUE)
  
  for (condition.use in names(avg.mat)) {
    CellChat::netVisual_circle(
      avg.mat[[condition.use]],
      weight.scale = TRUE,
      label.edge = TRUE,
      edge.weight.max = edge.max,
      edge.width.max = 12,
      title.name = paste0(title.prefix, " coarse mean number - ", condition.use)
    )
  }
}
```

# Compare the total number of interactions and interaction strength contact
Donor-averaged interaction number and strength

```{r}
plot_donor_average_number(object.list.contact.single, "Contact interaction number") +
  plot_donor_average_strength(object.list.contact.single, "Contact interaction strength")
```

Donor-averaged interaction number

```{r}
plot_donor_average_number(object.list.secreted.single, "Secreted interaction number")
```

Donor-averaged interaction number heatmap

```{r}
plot_average_count_heatmap(object.list.contact.single, "Contact donor-averaged interaction number")
```

secreted

Donor-averaged interaction number heatmap

```{r}
plot_average_count_heatmap(object.list.secreted.single, "Secreted donor-averaged interaction number")
```

# Pooled multi-sample interaction number and strength

Contact

```{r}
gg1_contact <- CellChat::compareInteractions(
  cellchat.contact,
  show.legend = FALSE,
  group = c(1, 2)
)

gg2_contact <- CellChat::compareInteractions(
  cellchat.contact,
  show.legend = FALSE,
  group = c(1, 2),
  measure = "weight"
)

gg1_contact + gg2_contact
```

Secreted

```{r}
gg1_secreted <- CellChat::compareInteractions(
  cellchat.secreted,
  show.legend = FALSE,
  group = c(1, 2)
)

gg2_secreted <- CellChat::compareInteractions(
  cellchat.secreted,
  show.legend = FALSE,
  group = c(1, 2),
  measure = "weight"
)

gg1_secreted + gg2_secreted
```


# Compare the major sources and targets in a 2D space (A) Identify cell populations with significant changes in sending or receiving signals Contact

> Note: **中等受影响**. Sender/receiver role is computed from pooled condition objects; interpret as pooled Control vs pooled Disease, not donor-average role.

```{r}
num.link.contact <- sapply(
  object.list.contact,
  function(x) {
    rowSums(x@net$count) +
      colSums(x@net$count) -
      diag(x@net$count)
  }
)

weight.MinMax.contact <- c(
  min(num.link.contact),
  max(num.link.contact)
)

gg.contact.role <- list()

for (i in seq_along(object.list.contact)) {
  gg.contact.role[[i]] <- SpatialCellChat::netAnalysis_signalingRole_scatter(
    object.list.contact[[i]],
    title = names(object.list.contact)[i],
    weight.MinMax = weight.MinMax.contact,
    x.measure = "outdeg",
    y.measure = "indeg"
  )
}

patchwork::wrap_plots(plots = gg.contact.role)
```

> Note: **中等受影响**. Sender/receiver role is computed from pooled condition objects; interpret as pooled Control vs pooled Disease, not donor-average role.

```{r}
num.link.secreted <- sapply(
  object.list.secreted,
  function(x) {
    rowSums(x@net$count) +
      colSums(x@net$count) -
      diag(x@net$count)
  }
)

weight.MinMax.secreted <- c(
  min(num.link.secreted),
  max(num.link.secreted)
)

gg.secreted.role <- list()

for (i in seq_along(object.list.secreted)) {
  gg.secreted.role[[i]] <- SpatialCellChat::netAnalysis_signalingRole_scatter(
    object.list.secreted[[i]],
    title = names(object.list.secreted)[i],
    weight.MinMax = weight.MinMax.secreted,
    x.measure = "outdeg",
    y.measure = "indeg"
  )
}

patchwork::wrap_plots(plots = gg.secreted.role)
```

# 1. Identify the signaling changes of specific cell populations contact

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg_eevt_contact <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.contact,
  idents.use = "eEVT"
)
```

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg_endo_contact <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.contact,
  idents.use = "Endothelial"
)
```

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg_evt_contact <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.contact,
  idents.use = "EVT"
)
```

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg_eevt_contact
```
CD80 immune cells? under stress?

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg_endo_contact
```

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg_evt_contact
```

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg2_eevt_secreted <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.secreted,
  idents.use = "eEVT"
)
```

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg2_endo_secreted <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.secreted,
  idents.use = "Endothelial"
)
```

> Note: **中等受影响**. Signaling change scatter uses pooled centrality differences; it is not donor-normalized.

```{r}
gg2_evt_secreted <- SpatialCellChat::netAnalysis_signalingChanges_scatter(
  cellchat.secreted,
  idents.use = "EVT"
)
```

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg2_eevt_secreted
```
negative value?

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg2_endo_secreted
```

> Note: **中等受影响**. This displays a pooled centrality-difference scatter; use only as pooled-condition evidence.

```{r}
gg2_evt_secreted
```

> Note: **不受影响**. Functional pathway clustering compares pathway communication patterns, not total interaction counts; describe it as pooled-condition pattern clustering.

```{r}
cellchat.contact <- SpatialCellChat::computeNetSimilarityPairwise(
  cellchat.contact,
  type = "functional"
)

cellchat.contact <- SpatialCellChat::netEmbedding(
  cellchat.contact,
  type = "functional"
)

cellchat.contact <- SpatialCellChat::netClustering(
  cellchat.contact,
  type = "functional"
)

SpatialCellChat::netVisual_embeddingPairwise(
  cellchat.contact,
  type = "functional",
  label.size = 3.5
)
```

secreted

> Note: **不受影响**. Functional pathway clustering compares pathway communication patterns, not total interaction counts; describe it as pooled-condition pattern clustering.

```{r}
cellchat.secreted <- SpatialCellChat::computeNetSimilarityPairwise(
  cellchat.secreted,
  type = "functional"
)

cellchat.secreted <- SpatialCellChat::netEmbedding(
  cellchat.secreted,
  type = "functional"
)

cellchat.secreted <- SpatialCellChat::netClustering(
  cellchat.secreted,
  type = "functional"
)

SpatialCellChat::netVisual_embeddingPairwise(
  cellchat.secreted,
  type = "functional",
  label.size = 3.5
)
```

# Identify signaling groups based on structure similarity contact

> Note: **不受影响**. Structural pathway clustering compares pathway network structure, not total interaction counts; describe it as pooled-condition pattern clustering.

```{r}
cellchat.contact <- SpatialCellChat::computeNetSimilarityPairwise(
  cellchat.contact,
  type = "structural"
)

cellchat.contact <- SpatialCellChat::netEmbedding(
  cellchat.contact,
  type = "structural"
)

cellchat.contact <- SpatialCellChat::netClustering(
  cellchat.contact,
  type = "structural"
)

SpatialCellChat::netVisual_embeddingPairwise(
  cellchat.contact,
  type = "structural",
  label.size = 3.5
)

SpatialCellChat::netVisual_embeddingPairwiseZoomIn(
  cellchat.contact,
  type = "structural",
  nCol = 2
)
```

secreted

> Note: **不受影响**. Structural pathway clustering compares pathway network structure, not total interaction counts; describe it as pooled-condition pattern clustering.

```{r}
cellchat.secreted <- SpatialCellChat::computeNetSimilarityPairwise(
  cellchat.secreted,
  type = "structural"
)

cellchat.secreted <- SpatialCellChat::netEmbedding(
  cellchat.secreted,
  type = "structural"
)

cellchat.secreted <- SpatialCellChat::netClustering(
  cellchat.secreted,
  type = "structural"
)

SpatialCellChat::netVisual_embeddingPairwise(
  cellchat.secreted,
  type = "structural",
  label.size = 3.5
)

SpatialCellChat::netVisual_embeddingPairwiseZoomIn(
  cellchat.secreted,
  type = "structural",
  nCol = 2
)
```


# Identify altered signaling with distinct interaction strength (A) Compare the overall information flow of each signaling pathway or ligand-receptor pair contact

> Note: **中等受影响**. rankNet compares pooled information flow; use donor-level pathway summaries for donor-normalized conclusions.

```{r}
gg1_contact_ranknet <- SpatialCellChat::rankNet(
  cellchat.contact,
  mode = "comparison",
  measure = "weight",
  sources.use = NULL,
  targets.use = NULL,
  stacked = TRUE,
  do.stat = TRUE
)

gg2_contact_ranknet <- SpatialCellChat::rankNet(
  cellchat.contact,
  mode = "comparison",
  measure = "weight",
  sources.use = NULL,
  targets.use = NULL,
  stacked = FALSE,
  do.stat = TRUE
)

gg1_contact_ranknet + gg2_contact_ranknet
```

> Note: **中等受影响**. rankNet compares pooled information flow; use donor-level pathway summaries for donor-normalized conclusions.

```{r}
gg1_secreted_ranknet <- SpatialCellChat::rankNet(
  cellchat.secreted,
  mode = "comparison",
  measure = "weight",
  sources.use = NULL,
  targets.use = NULL,
  stacked = TRUE,
  do.stat = TRUE
)

gg2_secreted_ranknet <- SpatialCellChat::rankNet(
  cellchat.secreted,
  mode = "comparison",
  measure = "weight",
  sources.use = NULL,
  targets.use = NULL,
  stacked = FALSE,
  do.stat = TRUE
)

gg1_secreted_ranknet + gg2_secreted_ranknet
```
Contact
```{r}
# Use the first two datasets, following the CellChat tutorial
i <- 1
j <- 2

# Combine all identified signaling pathways from the two datasets
pathway.union.contact <- union(
  object.list.contact[[i]]@netP$pathways,
  object.list.contact[[j]]@netP$pathways
)

# =========================
# Outgoing signaling
# =========================

ht_contact_out_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[i]],
  pattern = "outgoing",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[i],
  width = 5,
  height = 6
)

ht_contact_out_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[j]],
  pattern = "outgoing",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[j],
  width = 5,
  height = 6
)

ComplexHeatmap::draw(
  ht_contact_out_1 + ht_contact_out_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

```{r}
# =========================
# Incoming signaling
# =========================

ht_contact_in_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[i]],
  pattern = "incoming",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[i],
  width = 5,
  height = 6,
  color.heatmap = "GnBu"
)

ht_contact_in_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[j]],
  pattern = "incoming",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[j],
  width = 5,
  height = 6,
  color.heatmap = "GnBu"
)

ComplexHeatmap::draw(
  ht_contact_in_1 + ht_contact_in_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

```{r}
# =========================
# Overall signaling
# outgoing + incoming
# =========================

ht_contact_all_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[i]],
  pattern = "all",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[i],
  width = 5,
  height = 6,
  color.heatmap = "OrRd"
)

ht_contact_all_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.contact[[j]],
  pattern = "all",
  signaling = pathway.union.contact,
  title = names(object.list.contact)[j],
  width = 5,
  height = 6,
  color.heatmap = "OrRd"
)

ComplexHeatmap::draw(
  ht_contact_all_1 + ht_contact_all_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

Secreted

```{r}
i <- 1
j <- 2

pathway.union.secreted <- union(
  object.list.secreted[[i]]@netP$pathways,
  object.list.secreted[[j]]@netP$pathways
)

# =========================
# Outgoing signaling
# =========================

ht_secreted_out_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[i]],
  pattern = "outgoing",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[i],
  width = 5,
  height = 6
)

ht_secreted_out_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[j]],
  pattern = "outgoing",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[j],
  width = 5,
  height = 6
)

ComplexHeatmap::draw(
  ht_secreted_out_1 + ht_secreted_out_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

```{r}
# =========================
# Incoming signaling
# =========================

ht_secreted_in_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[i]],
  pattern = "incoming",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[i],
  width = 5,
  height = 6,
  color.heatmap = "GnBu"
)

ht_secreted_in_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[j]],
  pattern = "incoming",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[j],
  width = 5,
  height = 6,
  color.heatmap = "GnBu"
)

ComplexHeatmap::draw(
  ht_secreted_in_1 + ht_secreted_in_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

```{r}
# =========================
# Overall signaling
# outgoing + incoming
# =========================

ht_secreted_all_1 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[i]],
  pattern = "all",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[i],
  width = 5,
  height = 6,
  color.heatmap = "OrRd"
)

ht_secreted_all_2 <- SpatialCellChat::netAnalysis_signalingRole_heatmap(
  object.list.secreted[[j]],
  pattern = "all",
  signaling = pathway.union.secreted,
  title = names(object.list.secreted)[j],
  width = 5,
  height = 6,
  color.heatmap = "OrRd"
)

ComplexHeatmap::draw(
  ht_secreted_all_1 + ht_secreted_all_2,
  ht_gap = grid::unit(0.5, "cm")
)
```

# Part III: Identify the up-gulated and down-regulated signaling ligand-receptor pairs Identify dysfunctional signaling by comparing the communication probabities contact

> Note: **中等受影响**. Bubble plot is based on pooled LR probability differences; useful for candidate LR pairs, not donor-normalized conclusions.

```{r}
gg1 <- CellChat::netVisual_bubble(
  cellchat.contact,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  max.dataset = 2,
  title.name = "Increased signaling in Disease",
  angle.x = 45,
  remove.isolate = TRUE
)

gg2 <- CellChat::netVisual_bubble(
  cellchat.contact,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  max.dataset = 1,
  title.name = "Decreased signaling in Disease",
  angle.x = 45,
  remove.isolate = TRUE
)

gg1 + gg2
```
meaning of color

> Note: **中等受影响**. Bubble plot is based on pooled LR probability differences; useful for candidate LR pairs, not donor-normalized conclusions.

```{r}
gg1 <- CellChat::netVisual_bubble(
  cellchat.secreted,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  max.dataset = 2,
  title.name = "Increased signaling in Disease",
  angle.x = 45,
  remove.isolate = TRUE
)

gg2 <- CellChat::netVisual_bubble(
  cellchat.secreted,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  max.dataset = 1,
  title.name = "Decreased signaling in Disease",
  angle.x = 45,
  remove.isolate = TRUE
)

gg1 + gg2
```

# Identify dysfunctional signaling by using differential expression analysis contact

```{r}
pos.dataset <- "Disease"
features.name <- paste0(pos.dataset, ".merged")

cellchat.contact <- CellChat::identifyOverExpressedGenes(
  cellchat.contact,
  group.dataset = "datasets",
  pos.dataset = pos.dataset,
  features.name = features.name,
  only.pos = FALSE,
  thresh.pc = 0.1,
  thresh.fc = 0.05,
  thresh.p = 0.05
)
```

```{r}
# map DEA results onto inferred communications
net.contact <- CellChat::netMappingDEG(
  cellchat.contact,
  features.name = features.name
)

net.up.contact <- CellChat::subsetCommunication(
  cellchat.contact,
  net = net.contact,
  datasets = "Disease",
  ligand.logFC = 0.05,
  receptor.logFC = NULL
)

net.down.contact <- CellChat::subsetCommunication(
  cellchat.contact,
  net = net.contact,
  datasets = "Control",
  ligand.logFC = -0.05,
  receptor.logFC = NULL
)

gene.up.contact <- CellChat::extractGeneSubsetFromPair(
  net.up.contact,
  cellchat.contact
)

gene.down.contact <- CellChat::extractGeneSubsetFromPair(
  net.down.contact,
  cellchat.contact
)
```

secreted

```{r}
pos.dataset <- "Disease"
features.name <- paste0(pos.dataset, ".merged")

# perform differential expression analysis
cellchat.secreted <- CellChat::identifyOverExpressedGenes(
  cellchat.secreted,
  group.dataset = "datasets",
  pos.dataset = pos.dataset,
  features.name = features.name,
  only.pos = FALSE,
  thresh.pc = 0.1,
  thresh.fc = 0.05,
  thresh.p = 0.05
)
```

```{r}
# map DEG results onto inferred communications
net.secreted <- CellChat::netMappingDEG(
  cellchat.secreted,
  features.name = features.name
)

# LR pairs with upregulated ligands in Disease
net.up.secreted <- CellChat::subsetCommunication(
  cellchat.secreted,
  net = net.secreted,
  datasets = "Disease",
  ligand.logFC = 0.05,
  receptor.logFC = NULL
)

# LR pairs relatively downregulated in Disease
# i.e. higher in Control
net.down.secreted <- CellChat::subsetCommunication(
  cellchat.secreted,
  net = net.secreted,
  datasets = "Control",
  ligand.logFC = -0.05,
  receptor.logFC = NULL
)

# deconvolute signaling genes from LR pairs
gene.up.secreted <- CellChat::extractGeneSubsetFromPair(
  net.up.secreted,
  cellchat.secreted
)

gene.down.secreted <- CellChat::extractGeneSubsetFromPair(
  net.down.secreted,
  cellchat.secreted
)
```

Visualize the identified up-regulated and down-regulated signaling ligand-receptor pairs (A) Bubble plot contact

> Note: **中等受影响**. Up/down LR pairs are selected from pooled differences; use as candidates unless validated per donor.

```{r}
pairLR.use.up.contact <- net.up.contact[, "interaction_name", drop = FALSE]

gg1_contact <- CellChat::netVisual_bubble(
  cellchat.contact,
  pairLR.use = pairLR.use.up.contact,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  angle.x = 90,
  remove.isolate = TRUE,
  title.name = "Contact: up-regulated signaling in Disease"
)

pairLR.use.down.contact <- net.down.contact[, "interaction_name", drop = FALSE]

gg2_contact <- CellChat::netVisual_bubble(
  cellchat.contact,
  pairLR.use = pairLR.use.down.contact,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  angle.x = 90,
  remove.isolate = TRUE,
  title.name = "Contact: down-regulated signaling in Disease"
)

gg1_contact + gg2_contact
```

> Note: **中等受影响**. Up/down LR pairs are selected from pooled differences; use as candidates unless validated per donor.

```{r}
pairLR.use.up <- net.up.secreted[, "interaction_name", drop = FALSE]

gg1 <-  CellChat::netVisual_bubble(
  cellchat.secreted,
  pairLR.use = pairLR.use.up,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  angle.x = 90,
  remove.isolate = TRUE,
  title.name = "Up-regulated signaling in Disease"
)

pairLR.use.down <- net.down.secreted[, "interaction_name", drop = FALSE]

gg2 <-  CellChat::netVisual_bubble(
  cellchat.secreted,
  pairLR.use = pairLR.use.down,
  sources.use = c("eEVT", "Endothelial", "EVT"),
  targets.use = c("eEVT", "Endothelial", "EVT"),
  comparison = c(1, 2),
  angle.x = 90,
  remove.isolate = TRUE,
  title.name = "Down-regulated signaling in Disease"
)

gg1 + gg2
```

# Part V: Compare the signaling gene expression distribution between different datasets

> Note: **不受影响**. This is gene expression, not inferred interaction count; interpret as pooled expression distribution.

```{r}
# make sure dataset order is Control -> Disease
cellchat.contact@meta$datasets <- factor(
  cellchat.contact@meta$datasets,
  levels = c("Control", "Disease")
)

# Violin plot for NOTCH signaling genes
p_notch_expr <- CellChat::plotGeneExpression(
  cellchat.contact,
  signaling = "NOTCH",
  split.by = "datasets",
  colors.ggplot = TRUE,
  type = "violin"
)
```

> Note: **不受影响**. This displays the expression plot generated above, not inferred interaction count.

```{r}
p_notch_expr
```


# Circle plot of interaction number across datasets

Contact

Donor-averaged interaction number circle plot

```{r}
plot_average_count_circle(object.list.contact.single, "Contact")
```

Secreted

Donor-averaged interaction number circle plot

```{r}
plot_average_count_circle(object.list.secreted.single, "Secreted")
```

# Circle plot of coarse cell-type communication

Contact

Donor-averaged coarse interaction number circle plot

```{r}
plot_average_coarse_count_circle(object.list.contact.single, "Contact")
```

Secreted

Donor-averaged coarse interaction number circle plot

```{r}
plot_average_coarse_count_circle(object.list.secreted.single, "Secreted")
```

# Cell identity and key LR quality check

```{r}
obj_six <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_six_donors_Myometrium.rds"
)

DefaultAssay(obj_six) <- "Xenium"

identity_marker_panels <- list(
  `immune genes` = c(
    "CTLA4", "PDCD1",
    "NKG7", "KLRD1", "GNLY",
    "LST1", "TYROBP", "FCER1G",
    "LILRB2", "MRC1",
    "CEACAM8", "CCR3",
    "CCR4", "CCR5", "CCR7",
    "CXCR5", "CXCR6"
  ),
  `chemokine program` = c(
    "TNF",
    "CCL3", "CCL3L1",
    "CCL4", "CCL5",
    "CXCL13"
  ),
  `MHC-I / lineage check` = c(
    "HLA-G", "HLA-B",
    "CD8A", "CD8B",
    "PTPRC", "TRAC", "CD3D", "CD3E",
    "KRT7", "ITGA5", "MMP2",
    "PECAM1", "VWF"
  )
)

key_lr_gene_panels <- list(
  `contact LR hits` = c(
    "CEACAM1", "CEACAM6",
    "NCAM1",
    "JAG1", "NOTCH2",
    "CD80", "CTLA4",
    "CD274", "PDCD1LG2",
    "SELL",
    "ICOS",
    "LCK",
    "HLA-G", "HLA-B"
  ),
  `secreted LR hits` = c(
    "VEGFA", "FLT1", "KDR",
    "IL1B", "IL1R1",
    "IL2", "IL2RG",
    "IL4", "IL4R",
    "TNF", "TNFRSF1A",
    "CCL2", "CCL5",
    "CCR1", "CCR5",
    "CXCL8", "CXCL12", "CXCR4",
    "SPP1", "CD44",
    "CLEC11A", "KIT",
    "LEP", "LEPR"
  )
)

keep_present_markers <- function(obj, marker.panels) {
  panel.genes <- rownames(obj[["Xenium"]])
  markers.present <- lapply(
    marker.panels,
    function(x) intersect(x, panel.genes)
  )
  markers.present[lengths(markers.present) > 0]
}

make_identity_check <- function(obj, condition.use) {
  obj_sub <- subset(
    obj,
    subset = donor %in% names(sample.condition)[sample.condition == condition.use] &
      CellTypeManual.l2 %in% c("eEVT", "Endothelial", "EVT")
  )
  
  obj_sub$CellTypeManual.l2 <- factor(
    obj_sub$CellTypeManual.l2,
    levels = c("eEVT", "Endothelial", "EVT")
  )
  
  markers.present <- keep_present_markers(obj_sub, identity_marker_panels)
  
  Seurat::DotPlot(
    obj_sub,
    features = markers.present,
    group.by = "CellTypeManual.l2",
    assay = "Xenium",
    scale = TRUE,
    dot.scale = 6
  ) +
    Seurat::RotatedAxis() +
    labs(
      title = paste0(condition.use, ": eEVT, Endothelial and EVT identity check"),
      x = "Marker genes",
      y = "Cell identity"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      axis.text.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5),
      strip.text = element_text(size = 8)
    )
}

make_key_lr_check <- function(obj, condition.use) {
  obj_sub <- subset(
    obj,
    subset = donor %in% names(sample.condition)[sample.condition == condition.use] &
      CellTypeManual.l2 %in% c("eEVT", "Endothelial", "EVT")
  )
  
  obj_sub$CellTypeManual.l2 <- factor(
    obj_sub$CellTypeManual.l2,
    levels = c("eEVT", "Endothelial", "EVT")
  )
  
  markers.present <- keep_present_markers(obj_sub, key_lr_gene_panels)
  
  Seurat::DotPlot(
    obj_sub,
    features = markers.present,
    group.by = "CellTypeManual.l2",
    assay = "Xenium",
    scale = TRUE,
    dot.scale = 3.2,
    dot.min = 0.01
  ) +
    Seurat::RotatedAxis() +
    labs(
      title = paste0(condition.use, ": key LR gene check"),
      x = "Key LR genes",
      y = "Cell type"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 7),
      axis.text.y = element_text(size = 9),
      axis.title.x = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11),
      strip.text = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.title = element_text(size = 8),
      panel.spacing = unit(0.8, "lines")
    )
}
```

## Cell identity QC

```{r, fig.width=10, fig.height=5}
for (condition.use in c("Control", "Disease")) {
  print(make_identity_check(obj_six, condition.use))
}
```

## Key LR gene QC

```{r, fig.width=10, fig.height=5}
for (condition.use in c("Control", "Disease")) {
  print(make_key_lr_check(obj_six, condition.use))
}
```
