---
title: "Liana"
output: html_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(magrittr)
library(liana)
```
```{r}
show_resources()
```
Ligand receptor database: which ligand-receptor pairs can be recognize
```{r}
show_methods()
```
Scoring method: Whether specific ligand-receptor pairs are significant.

```{r}
liana_path <- system.file(package = "liana")

testdata <- readRDS(
  file.path(liana_path, "testdata", "input", "testdata.rds")
)
testdata %>% dplyr::glimpse()
```
Loading small example dataset. 90 cells, three cell types.

```{r}
# Run liana
liana_test <- liana_wrap(testdata)
# Liana returns a list of results, each element of which corresponds to a method
liana_test %>% dplyr::glimpse()
```
Five scoring method. 
Source: Source cell type, expressing ligand
Target: Target cell type, expressing receptor
Ligand
Receptor
ligand.prop: Proportion of cells expressing this ligand
receptor.prop: Proportion of cells expressing this receptor
ligand.expr:Average expression of ligand in the cell type
receptor.expr:Average expression of receptor in the cell type
Different scoring method will have different result.

```{r}
# We can aggregate these results into a tibble with consensus ranks
liana_test <- liana_test %>%
  liana_aggregate()
dplyr::glimpse(liana_test)
```
COmbine result from different scoring method to get a consensus result. 
The lower the aggregate_rank and mean_rank, the stronger the interaction. 

```{r}
liana_test %>%
  liana_dotplot(source_groups = c("B"),
                target_groups = c("NK", "CD8 T", "B"),
                ntop = 20)
```
Top 20 interactions listed on y-axis. From B cell to B, CD8 T, and NK cells.
Specificity: Is this interaction only in this pair of cell types?

```{r}
liana_trunc <- liana_test %>%
   # only keep interactions concordant between methods
  filter(aggregate_rank <= 0.01) # note that these pvals are already corrected

heat_freq(liana_trunc)
```
Count the number of interaction

```{r}
# Run liana re-implementations with the CellPhoneDB resource
complex_test <- liana_wrap(testdata,
                           method = c('natmi', 'sca', 'logfc'),
                           resource = c('CellPhoneDB'))
#> Warning in exec(output, ...): 3465 genes and/or 0 cells were removed as they had
#> no counts!

complex_test %>% liana_aggregate()
```

Liana can also deal with ligand-receptor complex with multiple subunit

```{r}
# define geometric mean
geometric_mean <- function(vec){exp(mean(log(vec)))}

# Overwrite default parameters by providing a list of parameters
liana_test <- liana_wrap(testdata,
                         method = c('cellphonedb', 'sca'),
                         resource = 'Consensus',
                         permutation.params = 
                           list(
                             nperms = 10 # here we run cpdb it only with 10 permutations
                             ),
                         complex_policy="geometric_mean"
                         )
#> Warning in exec(output, ...): 3465 genes and/or 0 cells were removed as they had
#> no counts!

# This returns a list of results for each method
liana_test %>% dplyr::glimpse()
```
We can modify parameters of liana, such as nperms to increase permutations. Here we use 10 only because we are doing a tuturial so we want to save time.
