---
title: "Neighbour_analysis_exploration.html"
output: html_document
date: "2026-05-31"
knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_dir = "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output",
      output_file = "Neighbour_analysis_exploration.html"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Matrix)
library(dplyr)
library(tidyr)
library(tibble)
library(Seurat)
library(knitr)
library(ggplot2)
library(RANN)       # nn2() - rapid nearest-neighbour computation
library(patchwork)
library(liana)

```
liana
NATMI: uses ligand–receptor expression plus cell-pair specificity, asking whether the interaction is relatively specific to the source–target cell pair.
Connectome: uses scaled/specificity-adjusted expression, prioritising interactions where ligand and receptor are relatively enriched in the relevant cell types.
logFC: checks whether the ligand is source-specific and the receptor is target-specific compared with other cell types.
SCA / SingleCellSignalR: mainly reflects ligand–receptor expression magnitude using a regularised LRscore.
CellPhoneDB: combines ligand–receptor mean expression with permutation-based support for whether the interaction is stronger than expected by chance.
```{r}
obj_control_fvb <- readRDS(file = "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_donor_FVB_Myometrium.rds")
obj_disease_fam <- readRDS(file = "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_donor_FAM_Myometrium.rds")
objs <- list(
  "Control FVB" = obj_control_fvb,
  "Disease FAM" = obj_disease_fam
)
```

```{r}
panel_genes <- rownames(obj_control_fvb[["Xenium"]])

head(panel_genes)
```

```{r}
expr_counts <- GetAssayData(
  obj_control_fvb,
  assay = "Xenium",
  layer = "counts"
)

expr_counts[1:5, 1:5]
```

```{r}
# Use Xenium assay
DefaultAssay(obj_control_fvb) <- "Xenium"
DefaultAssay(obj_disease_fam) <- "Xenium"

# Set cell type as active identity
Idents(obj_control_fvb) <- "CellTypeManual.l2"
Idents(obj_disease_fam) <- "CellTypeManual.l2"

# Check identities
table(Idents(obj_control_fvb))
table(Idents(obj_disease_fam))
```
```{r}
obj_control_fvb <- NormalizeData(
  obj_control_fvb,
  assay = "Xenium",
  verbose = FALSE
)

obj_disease_fam <- NormalizeData(
  obj_disease_fam,
  assay = "Xenium",
  verbose = FALSE
)
```

```{r}
# Control FVB
liana_control <- liana_wrap(obj_control_fvb)

# Disease FAM
liana_disease <- liana_wrap(obj_disease_fam)
```
```{r}
liana_control <- liana_control %>%
  liana_aggregate()

liana_disease <- liana_disease %>%
  liana_aggregate()
```

```{r}
glimpse(liana_control)
glimpse(liana_disease)
```
```{r}
liana_control <- liana_control %>%
  mutate(dataset = "Control FVB", .before = 1)

liana_disease <- liana_disease %>%
  mutate(dataset = "Disease FAM", .before = 1)

liana_all <- bind_rows(
  liana_control,
  liana_disease
)
```

```{r}
liana_evt_endo <- liana_all %>%
  filter(
    (source == "eEVT" & target == "Endothelial") |
      (source == "Endothelial" & target == "eEVT")
  ) %>%
  arrange(dataset, source, target, aggregate_rank) %>%
  mutate(
    direction = paste(source, target, sep = " -> "),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  )
liana_evt_endo_top20 <- liana_evt_endo %>%
  group_by(dataset, direction) %>%
  slice_head(n = 20) %>%
  ungroup() %>%
  select(
    dataset,
    direction,
    lr_complex,
    ligand.complex,
    receptor.complex,
    aggregate_rank,
    mean_rank,
    everything()
  )

liana_evt_endo_top20
```
Top 20 in both dataset and both direction

```{r}
control_liana_trunc <- liana_control %>%
  filter(aggregate_rank <= 0.05) %>%
  arrange(aggregate_rank)
heat_freq(control_liana_trunc)

disease_liana_trunc <- liana_disease %>%
  filter(aggregate_rank <= 0.05)
heat_freq(disease_liana_trunc)
```


```{r}
sig_cutoff <- 0.05

# 1. First identify LR pairs that are significant in either dataset
evt_to_endo_sig_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "Endothelial",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(ligand.complex, receptor.complex, lr_complex)

# 2. Pull full LIANA results for those pairs from both datasets, not only significant rows
evt_to_endo_full_for_sig_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "Endothelial"
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  semi_join(
    evt_to_endo_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex")
  ) %>%
  group_by(dataset_clean, ligand.complex, receptor.complex, lr_complex) %>%
  slice_min(aggregate_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    significant = if_else(aggregate_rank <= sig_cutoff, 1L, 0L)
  ) %>%
  select(
    dataset_clean,
    ligand.complex,
    receptor.complex,
    lr_complex,
    aggregate_rank,
    mean_rank,
    significant
  )

# 3. Make comparison table
evt_to_endo_compare <- expand_grid(
  evt_to_endo_sig_pairs,
  dataset_clean = c("control", "disease")
) %>%
  left_join(
    evt_to_endo_full_for_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex", "dataset_clean")
  ) %>%
  mutate(
    significant = replace_na(significant, 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = c(significant, aggregate_rank, mean_rank),
    names_glue = "{dataset_clean}_{.value}"
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    lr_complex
  )

evt_to_endo_compare
```
CLEC11A → KIT
CXCL12 → CXCR4
ICAM1 → ITGAX
CDH5 → CDH5
JAG1 → NOTCH3 
LEP → LEPR 

```{r}
sig_cutoff <- 0.05

# 1. Get union of significant eEVT -> Endothelial LR pairs from both datasets
evt_to_endo_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "Endothelial",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(dataset_clean, ligand.complex, receptor.complex, lr_complex)

# 2. Mark whether each LR pair was significant in each dataset
# Important: use nesting() so ligand/receptor pairs are not mixed up
evt_to_endo_presence <- evt_to_endo_pairs %>%
  mutate(significant = 1L) %>%
  complete(
    nesting(ligand.complex, receptor.complex, lr_complex),
    dataset_clean = c("control", "disease"),
    fill = list(significant = 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = significant,
    names_glue = "{dataset_clean}_significant"
  )
# 3. Get cell IDs and normalized data
expr_control_norm <- GetAssayData(
  obj_control_fvb,
  assay = "Xenium",
  layer = "data"
)

expr_disease_norm <- GetAssayData(
  obj_disease_fam,
  assay = "Xenium",
  layer = "data"
)

expr_control_counts <- GetAssayData(
  obj_control_fvb,
  assay = "Xenium",
  layer = "counts"
)

expr_disease_counts <- GetAssayData(
  obj_disease_fam,
  assay = "Xenium",
  layer = "counts"
)

# make sure expression columns match metadata rows
control_cells <- intersect(colnames(expr_control_norm), rownames(obj_control_fvb@meta.data))
disease_cells <- intersect(colnames(expr_disease_norm), rownames(obj_disease_fam@meta.data))

expr_control_norm <- expr_control_norm[, control_cells, drop = FALSE]
expr_control_counts <- expr_control_counts[, control_cells, drop = FALSE]

expr_disease_norm <- expr_disease_norm[, disease_cells, drop = FALSE]
expr_disease_counts <- expr_disease_counts[, disease_cells, drop = FALSE]
meta_control <- obj_control_fvb@meta.data
meta_disease <- obj_disease_fam@meta.data

control_cell_type <- meta_control[colnames(expr_control_norm), "CellTypeManual.l2"]
disease_cell_type <- meta_disease[colnames(expr_disease_norm), "CellTypeManual.l2"]

control_eevt_cells <- colnames(expr_control_norm)[control_cell_type == "eEVT"]
control_endo_cells <- colnames(expr_control_norm)[control_cell_type == "Endothelial"]

disease_eevt_cells <- colnames(expr_disease_norm)[disease_cell_type == "eEVT"]
disease_endo_cells <- colnames(expr_disease_norm)[disease_cell_type == "Endothelial"]

# 4. Calculate expression support for each LR pair
lr_expression_compare <- lapply(seq_len(nrow(evt_to_endo_presence)), function(i) {
  
  ligand <- evt_to_endo_presence$ligand.complex[i]
  receptor <- evt_to_endo_presence$receptor.complex[i]
  lr_complex <- evt_to_endo_presence$lr_complex[i]
  
  ligand_genes <- unlist(strsplit(ligand, "_"))
  receptor_genes <- unlist(strsplit(receptor, "_"))
  
  ligand_genes_control <- intersect(ligand_genes, rownames(expr_control_norm))
  receptor_genes_control <- intersect(receptor_genes, rownames(expr_control_norm))
  
  ligand_genes_disease <- intersect(ligand_genes, rownames(expr_disease_norm))
  receptor_genes_disease <- intersect(receptor_genes, rownames(expr_disease_norm))
  
  control_ligand_expr <- if (length(ligand_genes_control) > 0) {
    colMeans(expr_control_norm[ligand_genes_control, control_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(control_eevt_cells))
  }
  
  control_ligand_detected <- if (length(ligand_genes_control) > 0) {
    colSums(expr_control_counts[ligand_genes_control, control_eevt_cells, drop = FALSE] > 0) ==
      length(ligand_genes_control)
  } else {
    rep(FALSE, length(control_eevt_cells))
  }
  
  control_receptor_expr <- if (length(receptor_genes_control) > 0) {
    colMeans(expr_control_norm[receptor_genes_control, control_endo_cells, drop = FALSE])
  } else {
    rep(0, length(control_endo_cells))
  }
  
  control_receptor_detected <- if (length(receptor_genes_control) > 0) {
    colSums(expr_control_counts[receptor_genes_control, control_endo_cells, drop = FALSE] > 0) ==
      length(receptor_genes_control)
  } else {
    rep(FALSE, length(control_endo_cells))
  }
  
  disease_ligand_expr <- if (length(ligand_genes_disease) > 0) {
    colMeans(expr_disease_norm[ligand_genes_disease, disease_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(disease_eevt_cells))
  }
  
  disease_ligand_detected <- if (length(ligand_genes_disease) > 0) {
    colSums(expr_disease_counts[ligand_genes_disease, disease_eevt_cells, drop = FALSE] > 0) ==
      length(ligand_genes_disease)
  } else {
    rep(FALSE, length(disease_eevt_cells))
  }
  
  disease_receptor_expr <- if (length(receptor_genes_disease) > 0) {
    colMeans(expr_disease_norm[receptor_genes_disease, disease_endo_cells, drop = FALSE])
  } else {
    rep(0, length(disease_endo_cells))
  }
  
  disease_receptor_detected <- if (length(receptor_genes_disease) > 0) {
    colSums(expr_disease_counts[receptor_genes_disease, disease_endo_cells, drop = FALSE] > 0) ==
      length(receptor_genes_disease)
  } else {
    rep(FALSE, length(disease_endo_cells))
  }
  
  data.frame(
    lr_complex = lr_complex,
    ligand.complex = ligand,
    receptor.complex = receptor,
    
    control_ligand_expr_eEVT = mean(control_ligand_expr),
    disease_ligand_expr_eEVT = mean(disease_ligand_expr),
    
    control_receptor_expr_Endothelial = mean(control_receptor_expr),
    disease_receptor_expr_Endothelial = mean(disease_receptor_expr),
    
    control_ligand_detected_fraction_eEVT = mean(control_ligand_detected),
    disease_ligand_detected_fraction_eEVT = mean(disease_ligand_detected),
    
    control_receptor_detected_fraction_Endothelial = mean(control_receptor_detected),
    disease_receptor_detected_fraction_Endothelial = mean(disease_receptor_detected)
  )
}) %>%
  bind_rows()

# 5. Join expression data back to original presence table
evt_to_endo_expression_table <- evt_to_endo_presence %>%
  left_join(
    lr_expression_compare,
    by = c("lr_complex", "ligand.complex", "receptor.complex")
  ) %>%
  mutate(
    control_pair_expr_score = sqrt(
      control_ligand_expr_eEVT * control_receptor_expr_Endothelial
    ),
    disease_pair_expr_score = sqrt(
      disease_ligand_expr_eEVT * disease_receptor_expr_Endothelial
    ),
    
    control_pair_detected_score = sqrt(
      control_ligand_detected_fraction_eEVT *
        control_receptor_detected_fraction_Endothelial
    ),
    disease_pair_detected_score = sqrt(
      disease_ligand_detected_fraction_eEVT *
        disease_receptor_detected_fraction_Endothelial
    ),
    
    pair_expr_log2FC_disease_vs_control = log2(
      (disease_pair_expr_score + 1e-6) /
        (control_pair_expr_score + 1e-6)
    )
  ) %>%
  select(
    lr_complex,
    control_significant,
    disease_significant,
    
    control_ligand_expr_eEVT,
    disease_ligand_expr_eEVT,
    control_receptor_expr_Endothelial,
    disease_receptor_expr_Endothelial,
    control_pair_expr_score,
    disease_pair_expr_score,
    pair_expr_log2FC_disease_vs_control,
    
    control_ligand_detected_fraction_eEVT,
    disease_ligand_detected_fraction_eEVT,
    control_receptor_detected_fraction_Endothelial,
    disease_receptor_detected_fraction_Endothelial,
    control_pair_detected_score,
    disease_pair_detected_score
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    desc(control_pair_expr_score + disease_pair_expr_score),
    lr_complex
  )

evt_to_endo_expression_table
```
Ligand goes down, receptor goes up to compensate? Stress?
FN1 -> CD44,SPP1 -> CD44 are for adhesion. Increased adhesion in disease?
JAG1->NOTCH1 and LEP->LEP higher in healthy. For remodeling etc.
JAG1->NOTCH3 similar score. However, lower ligand higher receptor expression. Penalty for significant level-> larger than 0.05. May suggested compensation/stress?


```{r}
selected_pairs <- c(
  "FN1 -> CD44",
  "SPP1 -> CD44",
  "JAG1 -> NOTCH1",
  "JAG1 -> NOTCH3",
  "LEP -> LEPR"
)

liana_evt_endo_selected <- liana_all %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  filter(
    source == "eEVT",
    target == "Endothelial",
    lr_complex %in% selected_pairs
  ) %>%
  mutate(
    lr_complex = factor(lr_complex, levels = selected_pairs)
  )

p_control <- liana_evt_endo_selected %>%
  filter(dataset == "Control FVB") %>%
  arrange(lr_complex) %>%
  liana_dotplot(
    source_groups = "eEVT",
    target_groups = "Endothelial",
    ntop = length(selected_pairs),
    magnitude = "sca.LRscore",
    specificity = "natmi.edge_specificity",
    colour.label = "Expression level",
    size.label = "Interaction specificity"
  ) +
  ggtitle("Control FVB: eEVT -> Endothelial") +
  guides(
    size = guide_legend(order = 1),
    colour = guide_colourbar(order = 2)
  )

p_disease <- liana_evt_endo_selected %>%
  filter(dataset == "Disease FAM") %>%
  arrange(lr_complex) %>%
  liana_dotplot(
    source_groups = "eEVT",
    target_groups = "Endothelial",
    ntop = length(selected_pairs),
    magnitude = "sca.LRscore",
    specificity = "natmi.edge_specificity",
    colour.label = "Expression level",
    size.label = "Interaction specificity"
  ) +
  ggtitle("Disease FAM: eEVT -> Endothelial") +
  guides(
    size = guide_legend(order = 1),
    colour = guide_colourbar(order = 2)
  )

p_control
p_disease
```
```{r}
selected_pairs <- c(
  "FN1 -> CD44",
  "SPP1 -> CD44",
  "JAG1 -> NOTCH1",
  "JAG1 -> NOTCH3",
  "LEP -> LEPR"
)

# 1. Convert your wide expression table into long format for plotting
bar_plot_data <- bind_rows(
  evt_to_endo_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Control FVB",
      significant = control_significant,
      expression_score = control_pair_expr_score,
      detected_score = control_pair_detected_score
    ),
  
  evt_to_endo_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Disease FAM",
      significant = disease_significant,
      expression_score = disease_pair_expr_score,
      detected_score = disease_pair_detected_score
    )
) %>%
  filter(lr_complex %in% selected_pairs) %>%   # 如果想画全部 significant pairs，就删掉这一行
  mutate(
    dataset_clean = factor(
      dataset_clean,
      levels = c("Control FVB", "Disease FAM")
    ),
    lr_complex = factor(
      lr_complex,
      levels = selected_pairs
    ),
    detected_label = sprintf("%.2f", detected_score),
    significant_label = ifelse(significant == 1, "*", "")
  )
ggplot(
  bar_plot_data,
  aes(
    x = lr_complex,
    y = expression_score,
    fill = dataset_clean
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = detected_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "eEVT -> Endothelial expression support for selected LR pairs",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

```{r}
# 1. Literature-defined candidate LR pairs
selected_lr_manual <- tibble::tibble(
  ligand = c("CLEC11A", "CXCL12", "ICAM1", "CDH5","JAG1"),
  receptor = c("KIT", "CXCR4", "ITGAX", "CDH5","NOTCH1")
) %>%
  mutate(
    lr_complex = paste(ligand, receptor, sep = " -> "),
    pair_order = seq_len(n())
  )

# 2. Function to calculate manual expression support
calculate_manual_lr_expression <- function(
    obj,
    dataset_name,
    pairs,
    assay = "Xenium",
    celltype_col = "CellTypeManual.l2",
    source_type = "eEVT",
    target_type = "Endothelial"
) {
  
  expr_counts <- GetAssayData(obj, assay = assay, layer = "counts")
  expr_norm <- GetAssayData(obj, assay = assay, layer = "data")
  
  meta <- obj@meta.data
  cell_type <- meta[colnames(expr_counts), celltype_col]
  
  source_cells <- colnames(expr_counts)[cell_type == source_type]
  target_cells <- colnames(expr_counts)[cell_type == target_type]
  
  all_cell_types <- sort(unique(cell_type))
  
  get_complex_expr <- function(mat, genes, cells) {
    genes_use <- intersect(genes, rownames(mat))
    
    if (length(genes_use) == 0) {
      return(rep(NA_real_, length(cells)))
    }
    
    colMeans(mat[genes_use, cells, drop = FALSE])
  }
  
  get_complex_detected <- function(count_mat, genes, cells) {
    genes_use <- intersect(genes, rownames(count_mat))
    
    if (length(genes_use) == 0) {
      return(rep(NA, length(cells)))
    }
    
    colSums(count_mat[genes_use, cells, drop = FALSE] > 0) == length(genes_use)
  }
  
  get_mean_by_celltype <- function(mat, genes) {
    genes_use <- intersect(genes, rownames(mat))
    
    if (length(genes_use) == 0) {
      out <- rep(NA_real_, length(all_cell_types))
      names(out) <- all_cell_types
      return(out)
    }
    
    expr_vec <- colMeans(mat[genes_use, , drop = FALSE])
    mean_by_type <- tapply(expr_vec, cell_type, mean, na.rm = TRUE)
    
    out <- rep(0, length(all_cell_types))
    names(out) <- all_cell_types
    out[names(mean_by_type)] <- mean_by_type
    
    out
  }
  
  lapply(seq_len(nrow(pairs)), function(i) {
    
    ligand <- pairs$ligand[i]
    receptor <- pairs$receptor[i]
    lr_complex <- pairs$lr_complex[i]
    
    ligand_genes <- unlist(strsplit(ligand, "_"))
    receptor_genes <- unlist(strsplit(receptor, "_"))
    
    ligand_expr_source <- get_complex_expr(expr_norm, ligand_genes, source_cells)
    receptor_expr_target <- get_complex_expr(expr_norm, receptor_genes, target_cells)
    
    ligand_detected_source <- get_complex_detected(expr_counts, ligand_genes, source_cells)
    receptor_detected_target <- get_complex_detected(expr_counts, receptor_genes, target_cells)
    
    ligand_mean_source <- mean(ligand_expr_source, na.rm = TRUE)
    receptor_mean_target <- mean(receptor_expr_target, na.rm = TRUE)
    
    ligand_detected_fraction_source <- mean(ligand_detected_source, na.rm = TRUE)
    receptor_detected_fraction_target <- mean(receptor_detected_target, na.rm = TRUE)
    
    ligand_by_type <- get_mean_by_celltype(expr_norm, ligand_genes)
    receptor_by_type <- get_mean_by_celltype(expr_norm, receptor_genes)
    
    ligand_specificity_source <- ligand_by_type[source_type] / sum(ligand_by_type, na.rm = TRUE)
    receptor_specificity_target <- receptor_by_type[target_type] / sum(receptor_by_type, na.rm = TRUE)
    
    pair_expression_score <- sqrt(ligand_mean_source * receptor_mean_target)
    pair_detected_score <- sqrt(
      ligand_detected_fraction_source * receptor_detected_fraction_target
    )
    
    pair_specificity_score <- sqrt(
      ligand_specificity_source * receptor_specificity_target
    )
    
    data.frame(
      dataset = dataset_name,
      ligand = ligand,
      receptor = receptor,
      lr_complex = lr_complex,
      
      ligand_expr_eEVT = ligand_mean_source,
      receptor_expr_Endothelial = receptor_mean_target,
      pair_expression_score = pair_expression_score,
      
      ligand_detected_fraction_eEVT = ligand_detected_fraction_source,
      receptor_detected_fraction_Endothelial = receptor_detected_fraction_target,
      pair_detected_score = pair_detected_score,
      
      ligand_specificity_eEVT = ligand_specificity_source,
      receptor_specificity_Endothelial = receptor_specificity_target,
      pair_specificity_score = pair_specificity_score
    )
  }) %>%
    bind_rows()
}
manual_lr_expression <- bind_rows(
  calculate_manual_lr_expression(
    obj_control_fvb,
    dataset_name = "Control FVB",
    pairs = selected_lr_manual
  ),
  calculate_manual_lr_expression(
    obj_disease_fam,
    dataset_name = "Disease FAM",
    pairs = selected_lr_manual
  )
) %>%
  left_join(
    selected_lr_manual %>% select(lr_complex, pair_order),
    by = "lr_complex"
  ) %>%
  mutate(
    dataset = factor(dataset, levels = c("Control FVB", "Disease FAM")),
    lr_complex = factor(
      lr_complex,
      levels = selected_lr_manual$lr_complex
    ),
    specificity_label = ifelse(
      is.na(pair_specificity_score),
      "NA",
      sprintf("%.2f", pair_specificity_score)
    )
  )

manual_lr_expression
```

```{r}
ggplot(
  manual_lr_expression,
  aes(
    x = lr_complex,
    y = pair_expression_score,
    fill = dataset
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = specificity_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "Manual expression support for selected eEVT -> Endothelial LR pairs",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
Endo to eEVT
```{r}
sig_cutoff <- 0.05

# 1. First identify LR pairs that are significant in either dataset
endo_to_evt_sig_pairs <- liana_all %>%
  filter(
    source == "Endothelial",
    target == "eEVT",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(ligand.complex, receptor.complex, lr_complex)

# 2. Pull full LIANA results for those pairs from both datasets, not only significant rows
endo_to_evt_full_for_sig_pairs <- liana_all %>%
  filter(
    source == "Endothelial",
    target == "eEVT"
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  semi_join(
    endo_to_evt_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex")
  ) %>%
  group_by(dataset_clean, ligand.complex, receptor.complex, lr_complex) %>%
  slice_min(aggregate_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    significant = if_else(aggregate_rank <= sig_cutoff, 1L, 0L)
  ) %>%
  select(
    dataset_clean,
    ligand.complex,
    receptor.complex,
    lr_complex,
    aggregate_rank,
    mean_rank,
    significant
  )

# 3. Make comparison table
endo_to_evt_compare <- expand_grid(
  endo_to_evt_sig_pairs,
  dataset_clean = c("control", "disease")
) %>%
  left_join(
    endo_to_evt_full_for_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex", "dataset_clean")
  ) %>%
  mutate(
    significant = replace_na(significant, 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = c(significant, aggregate_rank, mean_rank),
    names_glue = "{dataset_clean}_{.value}"
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    lr_complex
  )

endo_to_evt_compare
```
```{r}
# 1. Get union of significant Endothelial -> eEVT LR pairs from both datasets
endo_to_evt_pairs <- liana_all %>%
  filter(
    source == "Endothelial",
    target == "eEVT",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(dataset_clean, ligand.complex, receptor.complex, lr_complex)

# 2. Mark whether each LR pair was significant in each dataset
endo_to_evt_presence <- endo_to_evt_pairs %>%
  mutate(significant = 1L) %>%
  complete(
    nesting(ligand.complex, receptor.complex, lr_complex),
    dataset_clean = c("control", "disease"),
    fill = list(significant = 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = significant,
    names_glue = "{dataset_clean}_significant"
  )

# 3. Calculate expression support for each LR pair
lr_expression_compare_endo_to_evt <- lapply(seq_len(nrow(endo_to_evt_presence)), function(i) {
  
  ligand <- endo_to_evt_presence$ligand.complex[i]
  receptor <- endo_to_evt_presence$receptor.complex[i]
  lr_complex <- endo_to_evt_presence$lr_complex[i]
  
  ligand_genes <- unlist(strsplit(ligand, "_"))
  receptor_genes <- unlist(strsplit(receptor, "_"))
  
  ligand_genes_control <- intersect(ligand_genes, rownames(expr_control_norm))
  receptor_genes_control <- intersect(receptor_genes, rownames(expr_control_norm))
  
  ligand_genes_disease <- intersect(ligand_genes, rownames(expr_disease_norm))
  receptor_genes_disease <- intersect(receptor_genes, rownames(expr_disease_norm))
  
  # Control: ligand in Endothelial, receptor in eEVT
  control_ligand_expr <- if (length(ligand_genes_control) > 0) {
    colMeans(expr_control_norm[ligand_genes_control, control_endo_cells, drop = FALSE])
  } else {
    rep(0, length(control_endo_cells))
  }
  
  control_ligand_detected <- if (length(ligand_genes_control) > 0) {
    colSums(expr_control_counts[ligand_genes_control, control_endo_cells, drop = FALSE] > 0) ==
      length(ligand_genes_control)
  } else {
    rep(FALSE, length(control_endo_cells))
  }
  
  control_receptor_expr <- if (length(receptor_genes_control) > 0) {
    colMeans(expr_control_norm[receptor_genes_control, control_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(control_eevt_cells))
  }
  
  control_receptor_detected <- if (length(receptor_genes_control) > 0) {
    colSums(expr_control_counts[receptor_genes_control, control_eevt_cells, drop = FALSE] > 0) ==
      length(receptor_genes_control)
  } else {
    rep(FALSE, length(control_eevt_cells))
  }
  
  # Disease: ligand in Endothelial, receptor in eEVT
  disease_ligand_expr <- if (length(ligand_genes_disease) > 0) {
    colMeans(expr_disease_norm[ligand_genes_disease, disease_endo_cells, drop = FALSE])
  } else {
    rep(0, length(disease_endo_cells))
  }
  
  disease_ligand_detected <- if (length(ligand_genes_disease) > 0) {
    colSums(expr_disease_counts[ligand_genes_disease, disease_endo_cells, drop = FALSE] > 0) ==
      length(ligand_genes_disease)
  } else {
    rep(FALSE, length(disease_endo_cells))
  }
  
  disease_receptor_expr <- if (length(receptor_genes_disease) > 0) {
    colMeans(expr_disease_norm[receptor_genes_disease, disease_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(disease_eevt_cells))
  }
  
  disease_receptor_detected <- if (length(receptor_genes_disease) > 0) {
    colSums(expr_disease_counts[receptor_genes_disease, disease_eevt_cells, drop = FALSE] > 0) ==
      length(receptor_genes_disease)
  } else {
    rep(FALSE, length(disease_eevt_cells))
  }
  
  data.frame(
    lr_complex = lr_complex,
    ligand.complex = ligand,
    receptor.complex = receptor,
    
    control_ligand_expr_Endothelial = mean(control_ligand_expr),
    disease_ligand_expr_Endothelial = mean(disease_ligand_expr),
    
    control_receptor_expr_eEVT = mean(control_receptor_expr),
    disease_receptor_expr_eEVT = mean(disease_receptor_expr),
    
    control_ligand_detected_fraction_Endothelial = mean(control_ligand_detected),
    disease_ligand_detected_fraction_Endothelial = mean(disease_ligand_detected),
    
    control_receptor_detected_fraction_eEVT = mean(control_receptor_detected),
    disease_receptor_detected_fraction_eEVT = mean(disease_receptor_detected)
  )
}) %>%
  bind_rows()

# 4. Join expression data back to original presence table
endo_to_evt_expression_table <- endo_to_evt_presence %>%
  left_join(
    lr_expression_compare_endo_to_evt,
    by = c("lr_complex", "ligand.complex", "receptor.complex")
  ) %>%
  mutate(
    control_pair_expr_score = sqrt(
      control_ligand_expr_Endothelial * control_receptor_expr_eEVT
    ),
    disease_pair_expr_score = sqrt(
      disease_ligand_expr_Endothelial * disease_receptor_expr_eEVT
    ),
    
    control_pair_detected_score = sqrt(
      control_ligand_detected_fraction_Endothelial *
        control_receptor_detected_fraction_eEVT
    ),
    disease_pair_detected_score = sqrt(
      disease_ligand_detected_fraction_Endothelial *
        disease_receptor_detected_fraction_eEVT
    ),
    
    pair_expr_log2FC_disease_vs_control = log2(
      (disease_pair_expr_score + 1e-6) /
        (control_pair_expr_score + 1e-6)
    )
  ) %>%
  select(
    lr_complex,
    control_significant,
    disease_significant,
    
    control_ligand_expr_Endothelial,
    disease_ligand_expr_Endothelial,
    control_receptor_expr_eEVT,
    disease_receptor_expr_eEVT,
    control_pair_expr_score,
    disease_pair_expr_score,
    pair_expr_log2FC_disease_vs_control,
    
    control_ligand_detected_fraction_Endothelial,
    disease_ligand_detected_fraction_Endothelial,
    control_receptor_detected_fraction_eEVT,
    disease_receptor_detected_fraction_eEVT,
    control_pair_detected_score,
    disease_pair_detected_score
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    desc(control_pair_expr_score + disease_pair_expr_score),
    lr_complex
  )

endo_to_evt_expression_table
```
```{r}
selected_pairs_endo_to_evt <- endo_to_evt_expression_table$lr_complex

liana_endo_to_evt_selected <- liana_all %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  filter(
    source == "Endothelial",
    target == "eEVT",
    lr_complex %in% selected_pairs_endo_to_evt
  ) %>%
  mutate(
    lr_complex = factor(lr_complex, levels = selected_pairs_endo_to_evt)
  )
# 1. Convert wide expression table into long format for plotting
bar_plot_data_endo_to_evt <- bind_rows(
  endo_to_evt_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Control FVB",
      significant = control_significant,
      expression_score = control_pair_expr_score,
      detected_score = control_pair_detected_score
    ),
  
  endo_to_evt_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Disease FAM",
      significant = disease_significant,
      expression_score = disease_pair_expr_score,
      detected_score = disease_pair_detected_score
    )
) %>%
  mutate(
    dataset_clean = factor(
      dataset_clean,
      levels = c("Control FVB", "Disease FAM")
    ),
    lr_complex = factor(
      lr_complex,
      levels = selected_pairs_endo_to_evt
    ),
    detected_label = sprintf("%.2f", detected_score)
  )

ggplot(
  bar_plot_data_endo_to_evt,
  aes(
    x = lr_complex,
    y = expression_score,
    fill = dataset_clean
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = detected_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "Endothelial -> eEVT expression support for selected LR pairs",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
JAG1 -> NOTCH2 disease only. Reverse direction of notch signalling?
TGFB1 -> SMAD3 
FN1 -> SDC1 Control only.Matrix remodelling
Many EGFR. May just because eEVT express a lot of EGFR?

```{r}
sig_cutoff <- 0.05

# 1. First identify LR pairs that are significant in either dataset
eevt_to_eevt_sig_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "eEVT",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(ligand.complex, receptor.complex, lr_complex)

# 2. Pull full LIANA results for those pairs from both datasets, not only significant rows
eevt_to_eevt_full_for_sig_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "eEVT"
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  semi_join(
    eevt_to_eevt_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex")
  ) %>%
  group_by(dataset_clean, ligand.complex, receptor.complex, lr_complex) %>%
  slice_min(aggregate_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    significant = if_else(aggregate_rank <= sig_cutoff, 1L, 0L)
  ) %>%
  select(
    dataset_clean,
    ligand.complex,
    receptor.complex,
    lr_complex,
    aggregate_rank,
    mean_rank,
    significant
  )

# 3. Make comparison table
eevt_to_eevt_compare <- expand_grid(
  eevt_to_eevt_sig_pairs,
  dataset_clean = c("control", "disease")
) %>%
  left_join(
    eevt_to_eevt_full_for_sig_pairs,
    by = c("ligand.complex", "receptor.complex", "lr_complex", "dataset_clean")
  ) %>%
  mutate(
    significant = replace_na(significant, 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = c(significant, aggregate_rank, mean_rank),
    names_glue = "{dataset_clean}_{.value}"
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    lr_complex
  )

eevt_to_eevt_compare
```
```{r}
# 1. Get union of significant eEVT -> eEVT LR pairs from both datasets
eevt_to_eevt_pairs <- liana_all %>%
  filter(
    source == "eEVT",
    target == "eEVT",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(dataset_clean, ligand.complex, receptor.complex, lr_complex)

# 2. Mark whether each LR pair was significant in each dataset
eevt_to_eevt_presence <- eevt_to_eevt_pairs %>%
  mutate(significant = 1L) %>%
  complete(
    nesting(ligand.complex, receptor.complex, lr_complex),
    dataset_clean = c("control", "disease"),
    fill = list(significant = 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = significant,
    names_glue = "{dataset_clean}_significant"
  )

# 3. Calculate expression support for each LR pair
lr_expression_compare_eevt_to_eevt <- lapply(seq_len(nrow(eevt_to_eevt_presence)), function(i) {
  
  ligand <- eevt_to_eevt_presence$ligand.complex[i]
  receptor <- eevt_to_eevt_presence$receptor.complex[i]
  lr_complex <- eevt_to_eevt_presence$lr_complex[i]
  
  ligand_genes <- unlist(strsplit(ligand, "_"))
  receptor_genes <- unlist(strsplit(receptor, "_"))
  
  ligand_genes_control <- intersect(ligand_genes, rownames(expr_control_norm))
  receptor_genes_control <- intersect(receptor_genes, rownames(expr_control_norm))
  
  ligand_genes_disease <- intersect(ligand_genes, rownames(expr_disease_norm))
  receptor_genes_disease <- intersect(receptor_genes, rownames(expr_disease_norm))
  
  # Control: ligand in eEVT, receptor in eEVT
  control_ligand_expr <- if (length(ligand_genes_control) > 0) {
    colMeans(expr_control_norm[ligand_genes_control, control_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(control_eevt_cells))
  }
  
  control_ligand_detected <- if (length(ligand_genes_control) > 0) {
    colSums(expr_control_counts[ligand_genes_control, control_eevt_cells, drop = FALSE] > 0) ==
      length(ligand_genes_control)
  } else {
    rep(FALSE, length(control_eevt_cells))
  }
  
  control_receptor_expr <- if (length(receptor_genes_control) > 0) {
    colMeans(expr_control_norm[receptor_genes_control, control_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(control_eevt_cells))
  }
  
  control_receptor_detected <- if (length(receptor_genes_control) > 0) {
    colSums(expr_control_counts[receptor_genes_control, control_eevt_cells, drop = FALSE] > 0) ==
      length(receptor_genes_control)
  } else {
    rep(FALSE, length(control_eevt_cells))
  }
  
  # Disease: ligand in eEVT, receptor in eEVT
  disease_ligand_expr <- if (length(ligand_genes_disease) > 0) {
    colMeans(expr_disease_norm[ligand_genes_disease, disease_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(disease_eevt_cells))
  }
  
  disease_ligand_detected <- if (length(ligand_genes_disease) > 0) {
    colSums(expr_disease_counts[ligand_genes_disease, disease_eevt_cells, drop = FALSE] > 0) ==
      length(ligand_genes_disease)
  } else {
    rep(FALSE, length(disease_eevt_cells))
  }
  
  disease_receptor_expr <- if (length(receptor_genes_disease) > 0) {
    colMeans(expr_disease_norm[receptor_genes_disease, disease_eevt_cells, drop = FALSE])
  } else {
    rep(0, length(disease_eevt_cells))
  }
  
  disease_receptor_detected <- if (length(receptor_genes_disease) > 0) {
    colSums(expr_disease_counts[receptor_genes_disease, disease_eevt_cells, drop = FALSE] > 0) ==
      length(receptor_genes_disease)
  } else {
    rep(FALSE, length(disease_eevt_cells))
  }
  
  data.frame(
    lr_complex = lr_complex,
    ligand.complex = ligand,
    receptor.complex = receptor,
    
    control_ligand_expr_eEVT = mean(control_ligand_expr),
    disease_ligand_expr_eEVT = mean(disease_ligand_expr),
    
    control_receptor_expr_eEVT = mean(control_receptor_expr),
    disease_receptor_expr_eEVT = mean(disease_receptor_expr),
    
    control_ligand_detected_fraction_eEVT = mean(control_ligand_detected),
    disease_ligand_detected_fraction_eEVT = mean(disease_ligand_detected),
    
    control_receptor_detected_fraction_eEVT = mean(control_receptor_detected),
    disease_receptor_detected_fraction_eEVT = mean(disease_receptor_detected)
  )
}) %>%
  bind_rows()

# 4. Join expression data back to original presence table
eevt_to_eevt_expression_table <- eevt_to_eevt_presence %>%
  left_join(
    lr_expression_compare_eevt_to_eevt,
    by = c("lr_complex", "ligand.complex", "receptor.complex")
  ) %>%
  mutate(
    control_pair_expr_score = sqrt(
      control_ligand_expr_eEVT * control_receptor_expr_eEVT
    ),
    disease_pair_expr_score = sqrt(
      disease_ligand_expr_eEVT * disease_receptor_expr_eEVT
    ),
    
    control_pair_detected_score = sqrt(
      control_ligand_detected_fraction_eEVT *
        control_receptor_detected_fraction_eEVT
    ),
    disease_pair_detected_score = sqrt(
      disease_ligand_detected_fraction_eEVT *
        disease_receptor_detected_fraction_eEVT
    ),
    
    pair_expr_log2FC_disease_vs_control = log2(
      (disease_pair_expr_score + 1e-6) /
        (control_pair_expr_score + 1e-6)
    )
  ) %>%
  select(
    lr_complex,
    control_significant,
    disease_significant,
    
    control_ligand_expr_eEVT,
    disease_ligand_expr_eEVT,
    control_receptor_expr_eEVT,
    disease_receptor_expr_eEVT,
    control_pair_expr_score,
    disease_pair_expr_score,
    pair_expr_log2FC_disease_vs_control,
    
    control_ligand_detected_fraction_eEVT,
    disease_ligand_detected_fraction_eEVT,
    control_receptor_detected_fraction_eEVT,
    disease_receptor_detected_fraction_eEVT,
    control_pair_detected_score,
    disease_pair_detected_score
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    desc(control_pair_expr_score + disease_pair_expr_score),
    lr_complex
  )

eevt_to_eevt_expression_table
```
```{r}
selected_pairs_eevt_to_eevt <- eevt_to_eevt_expression_table$lr_complex

liana_eevt_to_eevt_selected <- liana_all %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  filter(
    source == "eEVT",
    target == "eEVT",
    lr_complex %in% selected_pairs_eevt_to_eevt
  ) %>%
  mutate(
    lr_complex = factor(lr_complex, levels = selected_pairs_eevt_to_eevt)
  )
bar_plot_data_eevt_to_eevt <- bind_rows(
  eevt_to_eevt_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Control FVB",
      significant = control_significant,
      expression_score = control_pair_expr_score,
      detected_score = control_pair_detected_score
    ),
  
  eevt_to_eevt_expression_table %>%
    transmute(
      lr_complex,
      dataset_clean = "Disease FAM",
      significant = disease_significant,
      expression_score = disease_pair_expr_score,
      detected_score = disease_pair_detected_score
    )
) %>%
  mutate(
    dataset_clean = factor(
      dataset_clean,
      levels = c("Control FVB", "Disease FAM")
    ),
    lr_complex = factor(
      lr_complex,
      levels = selected_pairs_eevt_to_eevt
    ),
    detected_label = sprintf("%.2f", detected_score)
  )

ggplot(
  bar_plot_data_eevt_to_eevt,
  aes(
    x = lr_complex,
    y = expression_score,
    fill = dataset_clean
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = detected_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "eEVT -> eEVT expression support for selected LR pairs",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
FN1->NT5E / FN1->SDC1 self remodeling. High specificity
JAG1->NOTCH2
INHBA->SMAD3 remodelling
VCAN->EGFR disease only.

```{r}
make_participating_evt_endo_cells <- function(
    obj,
    dataset_name,
    k_neighbours = 3,
    celltype_col = "CellTypeManual.l2",
    region_col = "regions",
    x_col = "centroid_X_1",
    y_col = "centroid_X_2"
) {
  
  meta_raw <- obj@meta.data
  
  # avoid duplicated column-name problems
  colnames(meta_raw) <- make.unique(colnames(meta_raw))
  
  meta <- meta_raw %>%
    mutate(
      cell_barcode = rownames(meta_raw),
      dataset = dataset_name,
      cell_type = .data[[celltype_col]],
      cell_region = .data[[region_col]],
      x = .data[[x_col]],
      y = .data[[y_col]]
    ) %>%
    filter(
      !is.na(x),
      !is.na(y),
      !is.na(cell_type),
      !is.na(cell_region)
    )
  
  neighbour_result <- meta %>%
    group_by(cell_region) %>%
    group_modify(~ {
      
      region_data <- .x
      
      if (nrow(region_data) <= k_neighbours) {
        return(tibble())
      }
      
      coords <- as.matrix(region_data[, c("x", "y")])
      
      nn <- RANN::nn2(
        data = coords,
        query = coords,
        k = k_neighbours + 1
      )
      
      nn_ids <- nn$nn.idx[, -1, drop = FALSE]
      nn_dists <- nn$nn.dists[, -1, drop = FALSE]
      
      lapply(seq_len(nrow(region_data)), function(i) {
        tibble(
          cell_barcode = region_data$cell_barcode[i],
          cell_type = region_data$cell_type[i],
          nearest_rank = seq_len(k_neighbours),
          nearest_cell_barcode = region_data$cell_barcode[nn_ids[i, ]],
          nearest_cell_type = region_data$cell_type[nn_ids[i, ]],
          distance = nn_dists[i, ]
        )
      }) %>%
        bind_rows()
    }) %>%
    ungroup()
  
  participating_cells <- neighbour_result %>%
    filter(cell_type %in% c("eEVT", "Endothelial")) %>%
    group_by(cell_barcode, cell_type) %>%
    summarise(
      has_required_neighbour = case_when(
        cell_type[1] == "eEVT" ~ any(nearest_cell_type == "Endothelial"),
        cell_type[1] == "Endothelial" ~ any(nearest_cell_type == "eEVT"),
        TRUE ~ FALSE
      ),
      .groups = "drop"
    ) %>%
    filter(has_required_neighbour) %>%
    pull(cell_barcode)
  
  participating_meta <- meta %>%
    filter(cell_barcode %in% participating_cells)
  
  participating_obj <- subset(
    obj,
    cells = participating_cells
  )
  
  list(
    object = participating_obj,
    meta = participating_meta,
    neighbour_table = neighbour_result,
    participating_cell_ids = participating_cells
  )
}
```

```{r}
control_participating <- make_participating_evt_endo_cells(
  obj_control_fvb,
  dataset_name = "Control FVB",
  k_neighbours = 3
)

disease_participating <- make_participating_evt_endo_cells(
  obj_disease_fam,
  dataset_name = "Disease FAM",
  k_neighbours = 3
)
obj_control_fvb_evt_endo_k3 <- control_participating$object
obj_disease_fam_evt_endo_k3 <- disease_participating$object
```

```{r}
class(control_participating)
class(disease_participating)

table(obj_control_fvb_evt_endo_k3@meta.data$CellTypeManual.l2)
table(obj_disease_fam_evt_endo_k3@meta.data$CellTypeManual.l2)

control_participating$meta %>%
  count(cell_region, cell_type)

disease_participating$meta %>%
  count(cell_region, cell_type)
```
```{r}
# Use Xenium assay for k3-neighbour participating cells
DefaultAssay(obj_control_fvb_evt_endo_k3) <- "Xenium"
DefaultAssay(obj_disease_fam_evt_endo_k3) <- "Xenium"

# Set cell type as active identity
Idents(obj_control_fvb_evt_endo_k3) <- "CellTypeManual.l2"
Idents(obj_disease_fam_evt_endo_k3) <- "CellTypeManual.l2"

# Normalize the subsetted objects
obj_control_fvb_evt_endo_k3 <- NormalizeData(
  obj_control_fvb_evt_endo_k3,
  assay = "Xenium",
  verbose = FALSE
)

obj_disease_fam_evt_endo_k3 <- NormalizeData(
  obj_disease_fam_evt_endo_k3,
  assay = "Xenium",
  verbose = FALSE
)
```

```{r}
# Run LIANA on k3-neighbour participating cells only

liana_control_k3 <- liana_wrap(
  obj_control_fvb_evt_endo_k3
)

liana_disease_k3 <- liana_wrap(
  obj_disease_fam_evt_endo_k3
)
```
```{r}
# Aggregate LIANA results

liana_control_k3 <- liana_control_k3 %>%
  liana_aggregate()

liana_disease_k3 <- liana_disease_k3 %>%
  liana_aggregate()

# Add dataset labels and combine

liana_control_k3 <- liana_control_k3 %>%
  mutate(dataset = "Control FVB", .before = 1)

liana_disease_k3 <- liana_disease_k3 %>%
  mutate(dataset = "Disease FAM", .before = 1)

liana_all_k3 <- bind_rows(
  liana_control_k3,
  liana_disease_k3
)
```
```{r}
control_liana_trunc_k3 <- liana_control_k3 %>%
  filter(aggregate_rank <= 0.05) %>%
  arrange(aggregate_rank)

heat_freq(control_liana_trunc_k3)

disease_liana_trunc_k3 <- liana_disease_k3 %>%
  filter(aggregate_rank <= 0.05) %>%
  arrange(aggregate_rank)

heat_freq(disease_liana_trunc_k3)
```

```{r}
sig_cutoff <- 0.05

control_liana_trunc_k3 <- liana_control_k3 %>%
  filter(aggregate_rank <= sig_cutoff) %>%
  arrange(aggregate_rank)

disease_liana_trunc_k3 <- liana_disease_k3 %>%
  filter(aggregate_rank <= sig_cutoff) %>%
  arrange(aggregate_rank)

control_liana_trunc_k3
disease_liana_trunc_k3
```

```{r}
heat_freq(control_liana_trunc_k3)
heat_freq(disease_liana_trunc_k3)
```

```{r}
sig_cutoff <- 0.05

# 1. First identify eEVT -> Endothelial LR pairs significant in either k3 dataset
evt_to_endo_sig_pairs_k3 <- liana_all_k3 %>%
  filter(
    source == "eEVT",
    target == "Endothelial",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(ligand.complex, receptor.complex, lr_complex)

# 2. Pull full LIANA results for those pairs from both k3 datasets
evt_to_endo_full_for_sig_pairs_k3 <- liana_all_k3 %>%
  filter(
    source == "eEVT",
    target == "Endothelial"
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  semi_join(
    evt_to_endo_sig_pairs_k3,
    by = c("ligand.complex", "receptor.complex", "lr_complex")
  ) %>%
  group_by(dataset_clean, ligand.complex, receptor.complex, lr_complex) %>%
  slice_min(aggregate_rank, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    significant = if_else(aggregate_rank <= sig_cutoff, 1L, 0L)
  ) %>%
  select(
    dataset_clean,
    ligand.complex,
    receptor.complex,
    lr_complex,
    aggregate_rank,
    mean_rank,
    significant
  )

# 3. Make comparison table
evt_to_endo_compare_k3 <- expand_grid(
  evt_to_endo_sig_pairs_k3,
  dataset_clean = c("control", "disease")
) %>%
  left_join(
    evt_to_endo_full_for_sig_pairs_k3,
    by = c("ligand.complex", "receptor.complex", "lr_complex", "dataset_clean")
  ) %>%
  mutate(
    significant = replace_na(significant, 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = c(significant, aggregate_rank, mean_rank),
    names_glue = "{dataset_clean}_{.value}"
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    lr_complex
  )

evt_to_endo_compare_k3
```
```{r}
# 4. Get normalized data and raw counts from k3 subset objects
expr_control_norm_k3 <- GetAssayData(
  obj_control_fvb_evt_endo_k3,
  assay = "Xenium",
  layer = "data"
)

expr_disease_norm_k3 <- GetAssayData(
  obj_disease_fam_evt_endo_k3,
  assay = "Xenium",
  layer = "data"
)

expr_control_counts_k3 <- GetAssayData(
  obj_control_fvb_evt_endo_k3,
  assay = "Xenium",
  layer = "counts"
)

expr_disease_counts_k3 <- GetAssayData(
  obj_disease_fam_evt_endo_k3,
  assay = "Xenium",
  layer = "counts"
)

meta_control_k3 <- obj_control_fvb_evt_endo_k3@meta.data
meta_disease_k3 <- obj_disease_fam_evt_endo_k3@meta.data

control_cell_type_k3 <- meta_control_k3[colnames(expr_control_norm_k3), "CellTypeManual.l2"]
disease_cell_type_k3 <- meta_disease_k3[colnames(expr_disease_norm_k3), "CellTypeManual.l2"]

control_eevt_cells_k3 <- colnames(expr_control_norm_k3)[control_cell_type_k3 == "eEVT"]
control_endo_cells_k3 <- colnames(expr_control_norm_k3)[control_cell_type_k3 == "Endothelial"]

disease_eevt_cells_k3 <- colnames(expr_disease_norm_k3)[disease_cell_type_k3 == "eEVT"]
disease_endo_cells_k3 <- colnames(expr_disease_norm_k3)[disease_cell_type_k3 == "Endothelial"]

# 5. Presence table for k3 significant pairs
evt_to_endo_pairs_k3 <- liana_all_k3 %>%
  filter(
    source == "eEVT",
    target == "Endothelial",
    aggregate_rank <= sig_cutoff
  ) %>%
  mutate(
    dataset_clean = case_when(
      dataset == "Control FVB" ~ "control",
      dataset == "Disease FAM" ~ "disease",
      TRUE ~ dataset
    ),
    lr_complex = paste(ligand.complex, receptor.complex, sep = " -> ")
  ) %>%
  distinct(dataset_clean, ligand.complex, receptor.complex, lr_complex)

evt_to_endo_presence_k3 <- evt_to_endo_pairs_k3 %>%
  mutate(significant = 1L) %>%
  complete(
    nesting(ligand.complex, receptor.complex, lr_complex),
    dataset_clean = c("control", "disease"),
    fill = list(significant = 0L)
  ) %>%
  pivot_wider(
    names_from = dataset_clean,
    values_from = significant,
    names_glue = "{dataset_clean}_significant"
  )
# 6. Calculate expression support for k3 eEVT -> Endothelial pairs
lr_expression_compare_k3 <- lapply(seq_len(nrow(evt_to_endo_presence_k3)), function(i) {
  
  ligand <- evt_to_endo_presence_k3$ligand.complex[i]
  receptor <- evt_to_endo_presence_k3$receptor.complex[i]
  lr_complex <- evt_to_endo_presence_k3$lr_complex[i]
  
  ligand_genes <- unlist(strsplit(ligand, "_"))
  receptor_genes <- unlist(strsplit(receptor, "_"))
  
  ligand_genes_control <- intersect(ligand_genes, rownames(expr_control_norm_k3))
  receptor_genes_control <- intersect(receptor_genes, rownames(expr_control_norm_k3))
  
  ligand_genes_disease <- intersect(ligand_genes, rownames(expr_disease_norm_k3))
  receptor_genes_disease <- intersect(receptor_genes, rownames(expr_disease_norm_k3))
  
  control_ligand_expr <- if (length(ligand_genes_control) > 0) {
    colMeans(expr_control_norm_k3[ligand_genes_control, control_eevt_cells_k3, drop = FALSE])
  } else {
    rep(0, length(control_eevt_cells_k3))
  }
  
  control_ligand_detected <- if (length(ligand_genes_control) > 0) {
    colSums(expr_control_counts_k3[ligand_genes_control, control_eevt_cells_k3, drop = FALSE] > 0) ==
      length(ligand_genes_control)
  } else {
    rep(FALSE, length(control_eevt_cells_k3))
  }
  
  control_receptor_expr <- if (length(receptor_genes_control) > 0) {
    colMeans(expr_control_norm_k3[receptor_genes_control, control_endo_cells_k3, drop = FALSE])
  } else {
    rep(0, length(control_endo_cells_k3))
  }
  
  control_receptor_detected <- if (length(receptor_genes_control) > 0) {
    colSums(expr_control_counts_k3[receptor_genes_control, control_endo_cells_k3, drop = FALSE] > 0) ==
      length(receptor_genes_control)
  } else {
    rep(FALSE, length(control_endo_cells_k3))
  }
  
  disease_ligand_expr <- if (length(ligand_genes_disease) > 0) {
    colMeans(expr_disease_norm_k3[ligand_genes_disease, disease_eevt_cells_k3, drop = FALSE])
  } else {
    rep(0, length(disease_eevt_cells_k3))
  }
  
  disease_ligand_detected <- if (length(ligand_genes_disease) > 0) {
    colSums(expr_disease_counts_k3[ligand_genes_disease, disease_eevt_cells_k3, drop = FALSE] > 0) ==
      length(ligand_genes_disease)
  } else {
    rep(FALSE, length(disease_eevt_cells_k3))
  }
  
  disease_receptor_expr <- if (length(receptor_genes_disease) > 0) {
    colMeans(expr_disease_norm_k3[receptor_genes_disease, disease_endo_cells_k3, drop = FALSE])
  } else {
    rep(0, length(disease_endo_cells_k3))
  }
  
  disease_receptor_detected <- if (length(receptor_genes_disease) > 0) {
    colSums(expr_disease_counts_k3[receptor_genes_disease, disease_endo_cells_k3, drop = FALSE] > 0) ==
      length(receptor_genes_disease)
  } else {
    rep(FALSE, length(disease_endo_cells_k3))
  }
  
  data.frame(
    lr_complex = lr_complex,
    ligand.complex = ligand,
    receptor.complex = receptor,
    
    control_ligand_expr_eEVT = mean(control_ligand_expr),
    disease_ligand_expr_eEVT = mean(disease_ligand_expr),
    
    control_receptor_expr_Endothelial = mean(control_receptor_expr),
    disease_receptor_expr_Endothelial = mean(disease_receptor_expr),
    
    control_ligand_detected_fraction_eEVT = mean(control_ligand_detected),
    disease_ligand_detected_fraction_eEVT = mean(disease_ligand_detected),
    
    control_receptor_detected_fraction_Endothelial = mean(control_receptor_detected),
    disease_receptor_detected_fraction_Endothelial = mean(disease_receptor_detected)
  )
}) %>%
  bind_rows()

# 7. Make final k3 expression table
evt_to_endo_expression_table_k3 <- evt_to_endo_presence_k3 %>%
  left_join(
    lr_expression_compare_k3,
    by = c("lr_complex", "ligand.complex", "receptor.complex")
  ) %>%
  mutate(
    control_pair_expr_score = sqrt(
      control_ligand_expr_eEVT * control_receptor_expr_Endothelial
    ),
    disease_pair_expr_score = sqrt(
      disease_ligand_expr_eEVT * disease_receptor_expr_Endothelial
    ),
    
    control_pair_detected_score = sqrt(
      control_ligand_detected_fraction_eEVT *
        control_receptor_detected_fraction_Endothelial
    ),
    disease_pair_detected_score = sqrt(
      disease_ligand_detected_fraction_eEVT *
        disease_receptor_detected_fraction_Endothelial
    ),
    
    pair_expr_log2FC_disease_vs_control = log2(
      (disease_pair_expr_score + 1e-6) /
        (control_pair_expr_score + 1e-6)
    )
  ) %>%
  select(
    lr_complex,
    control_significant,
    disease_significant,
    
    control_ligand_expr_eEVT,
    disease_ligand_expr_eEVT,
    control_receptor_expr_Endothelial,
    disease_receptor_expr_Endothelial,
    control_pair_expr_score,
    disease_pair_expr_score,
    pair_expr_log2FC_disease_vs_control,
    
    control_ligand_detected_fraction_eEVT,
    disease_ligand_detected_fraction_eEVT,
    control_receptor_detected_fraction_Endothelial,
    disease_receptor_detected_fraction_Endothelial,
    control_pair_detected_score,
    disease_pair_detected_score
  ) %>%
  arrange(
    desc(control_significant + disease_significant),
    desc(control_pair_expr_score + disease_pair_expr_score),
    lr_complex
  )

evt_to_endo_expression_table_k3
```
```{r}
selected_pairs_k3 <- evt_to_endo_expression_table_k3$lr_complex

bar_plot_data_k3 <- bind_rows(
  evt_to_endo_expression_table_k3 %>%
    transmute(
      lr_complex,
      dataset_clean = "Control FVB",
      significant = control_significant,
      expression_score = control_pair_expr_score,
      detected_score = control_pair_detected_score
    ),
  
  evt_to_endo_expression_table_k3 %>%
    transmute(
      lr_complex,
      dataset_clean = "Disease FAM",
      significant = disease_significant,
      expression_score = disease_pair_expr_score,
      detected_score = disease_pair_detected_score
    )
) %>%
  mutate(
    dataset_clean = factor(
      dataset_clean,
      levels = c("Control FVB", "Disease FAM")
    ),
    lr_complex = factor(
      lr_complex,
      levels = selected_pairs_k3
    ),
    detected_label = sprintf("%.2f", detected_score)
  )

ggplot(
  bar_plot_data_k3,
  aes(
    x = lr_complex,
    y = expression_score,
    fill = dataset_clean
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = detected_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "k3-neighbour eEVT -> Endothelial expression support",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
```{r}
# 1. Literature-defined candidate LR pairs
selected_lr_manual_k3 <- tibble::tibble(
  ligand = c("CLEC11A", "CXCL12", "ICAM1", "CDH5", "JAG1"),
  receptor = c("KIT", "CXCR4", "ITGAX", "CDH5", "NOTCH1")
) %>%
  mutate(
    lr_complex = paste(ligand, receptor, sep = " -> "),
    pair_order = seq_len(n())
  )

# 2. Calculate manual expression support using k3 participating cells
manual_lr_expression_k3 <- bind_rows(
  calculate_manual_lr_expression(
    obj_control_fvb_evt_endo_k3,
    dataset_name = "Control FVB",
    pairs = selected_lr_manual_k3
  ),
  calculate_manual_lr_expression(
    obj_disease_fam_evt_endo_k3,
    dataset_name = "Disease FAM",
    pairs = selected_lr_manual_k3
  )
) %>%
  left_join(
    selected_lr_manual_k3 %>% select(lr_complex, pair_order),
    by = "lr_complex"
  ) %>%
  mutate(
    dataset = factor(dataset, levels = c("Control FVB", "Disease FAM")),
    lr_complex = factor(
      lr_complex,
      levels = selected_lr_manual_k3$lr_complex
    ),
    specificity_label = ifelse(
      is.na(pair_specificity_score),
      "NA",
      sprintf("%.2f", pair_specificity_score)
    )
  )

manual_lr_expression_k3
```

```{r}
ggplot(
  manual_lr_expression_k3,
  aes(
    x = lr_complex,
    y = pair_expression_score,
    fill = dataset
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    colour = "black"
  ) +
  geom_text(
    aes(label = specificity_label),
    position = position_dodge(width = 0.8),
    vjust = -0.4,
    size = 3
  ) +
  labs(
    title = "Manual expression support for selected eEVT -> Endothelial LR pairs in k3-neighbour cells",
    x = "Ligand-receptor pair",
    y = "Pair expression score",
    fill = "Dataset"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15))
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```
```{r}
expr_counts <- GetAssayData(obj_control_fvb, assay = "Xenium", layer = "counts")
expr_norm <- GetAssayData(obj_control_fvb, assay = "Xenium", layer = "data")

meta <- obj_control_fvb@meta.data
cell_type <- meta[colnames(expr_counts), "CellTypeManual.l2"]

cga_cell_data <- data.frame(
  cell = colnames(expr_counts),
  cell_type = cell_type,
  CGA_count = as.numeric(expr_counts["CGA", ]),
  CGA_expr = as.numeric(expr_norm["CGA", ])
)

cga_summary <- cga_cell_data %>%
  group_by(cell_type) %>%
  summarise(
    n_cells = n(),
    mean_CGA_expr = mean(CGA_expr),
    detected_cells = sum(CGA_count > 0),
    detected_fraction = mean(CGA_count > 0),
    max_CGA_count = max(CGA_count),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_CGA_expr))

cga_summary

# define endothelial-only CGA table
endo_cga_check <- cga_cell_data %>%
  filter(cell_type == "Endothelial") %>%
  rename(CGA = CGA_count)

endo_cga_check %>%
  summarise(
    q50 = quantile(CGA, 0.50),
    q75 = quantile(CGA, 0.75),
    q90 = quantile(CGA, 0.90),
    q95 = quantile(CGA, 0.95),
    q99 = quantile(CGA, 0.99),
    max = max(CGA)
  )
```
CGA check

