---
title: "Six Donor"
author:
- name: Zack Guo
  affiliation: UCL Institute of Child Health
  email: wegmzgu@ucl.ac.uk
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float: yes
    toc_depth: '5'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Chunks for calculation are not run.Result starts from line 618
# Load the library
```{r}
library(SpatialCellChat)
library(Matrix)
library(patchwork)
library(dplyr)
library(ggplot2)
library(Seurat)
options(stringsAsFactors = F)

# SpatialCellChat progressr workaround for R Markdown
future::plan("sequential")
```

# Read six sample RDS, where there are two control and six diseases
```{r}
obj_six <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_six_donors_Myometrium.rds"
)

# Check metadata columns
colnames(obj_six@meta.data)

# Check assay names
Assays(obj_six)

# Set Xenium as default assay
DefaultAssay(obj_six) <- "Xenium"

#Check number of cells from each donor
sample_col <- "donor"

#FVB and FVQ are control, the rest samples are disease
table(obj_six@meta.data[[sample_col]])
```

# Function for all calculation in SpatialCellChat
```{r, eval=FALSE}
run_contact_cellchat_one_sample <- function(dataset, signaling.type) {
  
  ###Part I: Data input & processing and initialization of Spatial CellChat object
  
  
  ##Load the data
  
  #Load RDS
  obj_control_fvb <- dataset
  
  # Set Xenium as the active assay
  DefaultAssay(obj_control_fvb) <- "Xenium"
  
  # Generate normalized log-expression data
  obj_control_fvb <- NormalizeData(
    obj_control_fvb,
    assay = "Xenium",
    normalization.method = "LogNormalize",
    scale.factor = 10000,
    verbose = FALSE
  )
  
  #Prepare spatial coordinates of cells
  spatial.locs_control_fvb <- obj_control_fvb@meta.data[
    colnames(obj_control_fvb),
    c("centroid_X_1", "centroid_X_2"),
    drop = FALSE
  ]
  
  #Change name of columns to x and y
  colnames(spatial.locs_control_fvb) <- c("x", "y")
  
  #Define Xenium spatial factor. The spatial factor is already 1 so the conversion factor is just 1.
  conversion.factor <- 1
  spot.size <- 10
  
  #$tol=5 means it allows a 5 μm tolerance when defining spatial proximity.
  spatial.factors_control_fvb <- list(
    ratio = conversion.factor,
    tol = spot.size/2
  )
  
  
  ##Create a Spatial CellChat object
  
  chat_control_fvb <- createSpatialCellChat(
    object = obj_control_fvb,                  # Xenium Seurat object
    group.by = "CellTypeManual.l2",            # cell-type annotation
    assay = "Xenium",                          # normalized Xenium assay
    datatype = "spatial",
    coordinates = spatial.locs_control_fvb,    # cell centroid coordinates
    spatial.factors = spatial.factors_control_fvb
  )
  
  # Inspect the SpatialCellChat object
  chat_control_fvb
  
  # Check cell identities used in the analysis. x_cent/y_cent means x/y coordinate of cell centroid.
  table(chat_control_fvb@idents)
  
  
  ##Set the ligand-receptor interaction database
  
  custom_lr_clec11a_kit <- data.frame(
    ligand = "CLEC11A",
    receptor = "KIT",
    
    # Assign an individual pathway name so that the pair can be
    # retained in pathway-level CellChat analysis
    pathway_name = "CLEC11A",
    
    interaction_name = "CLEC11A_KIT",
    interaction_name_2 = "CLEC11A - KIT",
    
    # CLEC11A is treated as a soluble ligand
    annotation = "Secreted Signaling",
    
    evidence = paste(
      "NicheNet-predicted EVT-to-artery interaction;",
      "Greenbaum et al., Nature 2023, Fig. 6h;",
      "doi:10.1038/s41586-023-06298-9"
    ),
    
    stringsAsFactors = FALSE
  )
  
  # Merge the custom pair with the built-in human CellChat database
  CellChatDB_custom <- SpatialCellChat::updateCellChatDB(
    db = custom_lr_clec11a_kit,
    merged = TRUE,
    species_target = "human"
  )
  
  CellChatDB_use <- SpatialCellChat::subsetDB(
    CellChatDB_custom,
    search = c(
      "Secreted Signaling",
      "ECM-Receptor",
      "Cell-Cell Contact"
    ),
    non_protein = FALSE
  )
  
  
  ##Pre-processing the expression data for cell-cell communication analysis
  
  # Only keep contact signalling
  # Only keep genes in LR database.
  # Retain genes detected in at least 10 cells
  # Normally only spatially variable genes will be retain, but for Xenium data that has relatively small panel, we skip this
  
  
  # Select the human Cell-Cell Contact database
  CellChatDB <- CellChatDB.human
  
  CellChatDB_contact <- subsetDB(
    CellChatDB,
    search = "Cell-Cell Contact",
    non_protein = FALSE
  )
  
  CellChatDB_secreted <- subsetDB(
    CellChatDB_custom,
    search = "Secreted Signaling",
    non_protein = FALSE
  )
  
  CellChatDB_selected <- switch(
    signaling.type,
    contact = CellChatDB_contact,
    secreted = CellChatDB_secreted
  )
  
  # Assign the selected database to the SpatialCellChat object
  chat_control_fvb@DB <- CellChatDB_selected
  
  
  # Subset expression data to signalling genes
  # This step is required even when using the whole database
  chat_control_fvb <- subsetData(chat_control_fvb)
  
  
  # Count how many cells express each signalling gene
  # Do this BEFORE preProcessing/ALRA imputation
  n_cells_expressed <- Matrix::rowSums(
    chat_control_fvb@data.signaling > 0
  )
  
  
  # Retain genes detected in at least 10 cells
  min_cells_gene <- 10
  
  features_small_panel <- names(
    n_cells_expressed[n_cells_expressed >= min_cells_gene]
  )
  
  
  # Run SpatialCellChat preprocessing
  # This performs ALRA preprocessing on the signalling-expression matrix to make it compatible for spatial cellchat function
  chat_control_fvb <- preProcessing(chat_control_fvb)
  
  
  # Manually assign the genes retained for the small Xenium panel
  # This replaces identifyOverExpressedGenes()
  chat_control_fvb@var.features[["features"]] <- features_small_panel
  
  chat_control_fvb@var.features[["features.info"]] <- data.frame(
    features = names(n_cells_expressed),
    n_cells_expressed = as.numeric(n_cells_expressed),
    selected = n_cells_expressed >= min_cells_gene,
    row.names = names(n_cells_expressed),
    stringsAsFactors = FALSE
  )
  
  
  # Identify available ligand-receptor interactions
  # variable.both = FALSE means either the ligand or receptor must pass
  # the >=10-cell filter, while all required genes must be in the panel
  chat_control_fvb <- identifyOverExpressedInteractions(
    chat_control_fvb,
    variable.both = FALSE
  )
  
  
  ###Part II: Inference of cell-cell communication network
  
  
  ##Compute the communication probability and infer cellular communication network at the individual cell level(L-R)
  
  # Check the retained ligand-receptor pairs
  nrow(chat_control_fvb@LR$LRsig)
  
  # Confirm that the current analysis contains only contact signalling
  unique(chat_control_fvb@LR$LRsig$annotation)
  
  # Compute individual-cell LR communication probabilities using spatial distance
  chat_control_fvb <- switch(
    signaling.type,
    
    contact = computeCommunProb(
      chat_control_fvb,
      distance.use = TRUE,
      scale.distance = 0.8,
      contact.dependent = TRUE,
      interaction.range = 250,
      contact.range = 20
    ),
    
    secreted = computeCommunProb(
      chat_control_fvb,
      distance.use = TRUE,
      scale.distance = 0.8,
      contact.dependent = FALSE,
      interaction.range = 250
    )
  )
  
  
  # Filtering
  
  # Filter weak or poorly supported individual-cell LR communication links
  chat_control_fvb <- filterProbability(
    chat_control_fvb
  )
  
  # Keep only communication events with sufficient cell-cell link support
  chat_control_fvb <- filterCommunication(
    chat_control_fvb,
    min.cells = NULL,
    min.links = 10,
    min.cells.sr = 10
  )
  
  
  ## Calculate group-level LR communication probability
  
  # Average filtered individual-cell LR probabilities into source-target cell-group LR probabilities
  chat_control_fvb <- computeAvgCommunProb(
    chat_control_fvb,
    avg.type = "avg",
    min.percent = 0.1,
    min.cells.sr = 10,
    do.permutation = TRUE,
    nboot = 100,
    seed.use = 1L
  )
  
  # Remove group-level LR communications involving cell groups with fewer than 10 cells
  chat_control_fvb <- filterCommunication(
    chat_control_fvb,
    min.cells = 10,
    min.links = NULL,
    min.cells.sr = NULL
  )
  
  
  ##Infer communication at signaling-pathway level at cell group level
  
  chat_control_fvb <- computeCommunProbPathway(
    chat_control_fvb,
    thresh = 0.05,
    do.group = TRUE,
    do.cell = FALSE
  )
  
  
  ##Calculate the aggregated communication network
  
  # Aggregate LR/pathway communication probabilities into cell-group-level count and weight networks
  chat_control_fvb <- aggregateNet(
    chat_control_fvb
  )
  
  
  ##Compute network centrality scores
  
  # Signalling role of each cell type. Sender? Receiver?
  
  chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
    chat_control_fvb,
    slot.name = "net",
    do.group = FALSE,
    degree.only = TRUE
  )
  
  chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
    chat_control_fvb,
    slot.name = "net",
    do.group = TRUE,
    degree.only = TRUE
  )
  
  try(
    chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
      chat_control_fvb,
      slot.name = "netP",
      do.group = TRUE,
      degree.only = TRUE
    ),
    silent = TRUE
  )
  
  
  ##Return final SpatialCellChat object
  
  return(chat_control_fvb)
}
```

# Function for saving RDS
```{r, eval=FALSE}
save_cellchat_one_sample <- function(chat_object, sample_id, signaling.type) {
  
  output_data_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/single_sample_cellchat"
  
  dir.create(
    output_data_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
  
  sample_clean <- gsub("[^A-Za-z0-9_\\-]", "_", sample_id)
  
  file_suffix <- switch(
    signaling.type,
    contact = "contactRange20",
    secreted = "secreted_range250"
  )
  
  output_file <- file.path(
    output_data_dir,
    paste0(
      "chat_",
      sample_clean,
      "_",
      file_suffix,
      "_final.rds"
    )
  )
  
  saveRDS(
    chat_object,
    output_file
  )
  
  return(output_file)
}
```

# Run for FCM sample
```{r, eval=FALSE}
obj_fcm <- subset(
  obj_six,
  subset = donor == "FCM"
)

chat_fcm_contact <- run_contact_cellchat_one_sample(
  obj_fcm,
  signaling.type = "contact"
)

chat_fcm_secreted <- run_contact_cellchat_one_sample(
  obj_fcm,
  signaling.type = "secreted"
)
```

# Save FCM contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fcm_contact,
  sample_id = "FCM",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fcm_secreted,
  sample_id = "FCM",
  signaling.type = "secreted"
)
```

# Run for FEP sample
```{r, eval=FALSE}
obj_fep <- subset(
  obj_six,
  subset = donor == "FEP"
)

chat_fep_contact <- run_contact_cellchat_one_sample(
  obj_fep,
  signaling.type = "contact"
)

chat_fep_secreted <- run_contact_cellchat_one_sample(
  obj_fep,
  signaling.type = "secreted"
)
```

# Save FEP contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fep_contact,
  sample_id = "FEP",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fep_secreted,
  sample_id = "FEP",
  signaling.type = "secreted"
)
```

# Run for FVQ sample
```{r, eval=FALSE}
obj_fvq <- subset(
  obj_six,
  subset = donor == "FVQ"
)

chat_fvq_contact <- run_contact_cellchat_one_sample(
  obj_fvq,
  signaling.type = "contact"
)

chat_fvq_secreted <- run_contact_cellchat_one_sample(
  obj_fvq,
  signaling.type = "secreted"
)
```

# Save FVQ contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fvq_contact,
  sample_id = "FVQ",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvq_secreted,
  sample_id = "FVQ",
  signaling.type = "secreted"
)
```

# Run for FVS sample
```{r, eval=FALSE}
obj_fvs <- subset(
  obj_six,
  subset = donor == "FVS"
)

chat_fvs_contact <- run_contact_cellchat_one_sample(
  obj_fvs,
  signaling.type = "contact"
)

chat_fvs_secreted <- run_contact_cellchat_one_sample(
  obj_fvs,
  signaling.type = "secreted"
)
```

# Save FVS contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fvs_contact,
  sample_id = "FVS",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvs_secreted,
  sample_id = "FVS",
  signaling.type = "secreted"
)
```
# Run for FVB sample
```{r, eval=FALSE}
obj_fvb <- subset(
  obj_six,
  subset = donor == "FVB"
)

chat_fvb_contact <- run_contact_cellchat_one_sample(
  obj_fvb,
  signaling.type = "contact"
)

chat_fvb_secreted <- run_contact_cellchat_one_sample(
  obj_fvb,
  signaling.type = "secreted"
)
```

# Save FVB contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fvb_contact,
  sample_id = "FVB",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvb_secreted,
  sample_id = "FVB",
  signaling.type = "secreted"
)
```

# Run for FAM sample
```{r, eval=FALSE}
obj_fam <- subset(
  obj_six,
  subset = donor == "FAM"
)

chat_fam_contact <- run_contact_cellchat_one_sample(
  obj_fam,
  signaling.type = "contact"
)

chat_fam_secreted <- run_contact_cellchat_one_sample(
  obj_fam,
  signaling.type = "secreted"
)
```

# Save FAM contact and secreted objects
```{r, eval=FALSE}
save_cellchat_one_sample(
  chat_object = chat_fam_contact,
  sample_id = "FAM",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fam_secreted,
  sample_id = "FAM",
  signaling.type = "secreted"
)
```

# Loading 12 RDS
```{r}
single_sample_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/single_sample_cellchat"

#Two control samples
chat_fvb_contact <- readRDS(
  file.path(single_sample_dir, "chat_FVB_contactRange20_final.rds")
)

chat_fvb_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FVB_secreted_range250_final.rds")
)

chat_fvq_contact <- readRDS(
  file.path(single_sample_dir, "chat_FVQ_contactRange20_final.rds")
)

chat_fvq_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FVQ_secreted_range250_final.rds")
)

#Four disease samples
chat_fam_contact <- readRDS(
  file.path(single_sample_dir, "chat_FAM_contactRange20_final.rds")
)

chat_fam_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FAM_secreted_range250_final.rds")
)

chat_fcm_contact <- readRDS(
  file.path(single_sample_dir, "chat_FCM_contactRange20_final.rds")
)

chat_fcm_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FCM_secreted_range250_final.rds")
)

chat_fep_contact <- readRDS(
  file.path(single_sample_dir, "chat_FEP_contactRange20_final.rds")
)

chat_fep_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FEP_secreted_range250_final.rds")
)

chat_fvs_contact <- readRDS(
  file.path(single_sample_dir, "chat_FVS_contactRange20_final.rds")
)

chat_fvs_secreted <- readRDS(
  file.path(single_sample_dir, "chat_FVS_secreted_range250_final.rds")
)
```

# Check cell type composition in each loaded SpatialCellChat object
```{r}
celltype_counts <- list(
  FVB = table(chat_fvb_contact@idents),
  FVQ = table(chat_fvq_contact@idents),
  FAM = table(chat_fam_contact@idents),
  FCM = table(chat_fcm_contact@idents),
  FEP = table(chat_fep_contact@idents),
  FVS = table(chat_fvs_contact@idents)
)

celltype_counts
```

# Spatial plot for six single-sample SpatialCellChat objects
```{r}
# Fixed colors for cell types
celltype_cols <- c(
  "eEVT" = "#E64B35",
  "Endothelial" = "#4DBBD5",
  "EVT" = "#00A087"
)

p_fvb <- SpatialCellChat::spatialDimPlot(
  chat_fvb_contact,
  point.size = 1.2
) +
  ggtitle("FVB") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fvq <- SpatialCellChat::spatialDimPlot(
  chat_fvq_contact,
  point.size = 1.2
) +
  ggtitle("FVQ") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fam <- SpatialCellChat::spatialDimPlot(
  chat_fam_contact,
  point.size = 1.2
) +
  ggtitle("FAM") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fcm <- SpatialCellChat::spatialDimPlot(
  chat_fcm_contact,
  point.size = 1.2
) +
  ggtitle("FCM") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fep <- SpatialCellChat::spatialDimPlot(
  chat_fep_contact,
  point.size = 1.2
) +
  ggtitle("FEP") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fvs <- SpatialCellChat::spatialDimPlot(
  chat_fvs_contact,
  point.size = 1.2
) +
  ggtitle("FVS") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_fvb
p_fvq
p_fam
p_fcm
p_fep
p_fvs
```
# Extract significant signalling
```{r}
extract_sig_lr_safe <- function(chat, sample_id, signaling_type, thresh = 0.01) {
  out <- tryCatch(
    {
      SpatialCellChat::subsetCommunication(
        chat,
        slot.name = "net",
        thresh = thresh
      ) %>%
        mutate(
          sample = sample_id,
          signaling_type = signaling_type,
          .before = 1
        ) %>%
        arrange(pval, desc(prob))
    },
    error = function(e) {
      message(
        "No significant LR inferred for ",
        sample_id,
        " / ",
        signaling_type,
        " at thresh = ",
        thresh
      )
      return(NULL)
    }
  )
  
  return(out)
}
```

# All contact sample
```{r}
object.list.contact.single <- list(
  FVB = chat_fvb_contact,
  FVQ = chat_fvq_contact,
  FAM = chat_fam_contact,
  FCM = chat_fcm_contact,
  FEP = chat_fep_contact,
  FVS = chat_fvs_contact
)

sig_lr_contact_all <- bind_rows(
  lapply(names(object.list.contact.single), function(sample_id) {
    extract_sig_lr_safe(
      chat = object.list.contact.single[[sample_id]],
      sample_id = sample_id,
      signaling_type = "contact",
      thresh = 0.01
    )
  })
)

sig_lr_contact_all
```

```{r}
object.list.secreted.single <- list(
  FVB = chat_fvb_secreted,
  FVQ = chat_fvq_secreted,
  FAM = chat_fam_secreted,
  FCM = chat_fcm_secreted,
  FEP = chat_fep_secreted,
  FVS = chat_fvs_secreted
)

sig_lr_secreted_all <- bind_rows(
  lapply(names(object.list.secreted.single), function(sample_id) {
    extract_sig_lr_safe(
      chat = object.list.secreted.single[[sample_id]],
      sample_id = sample_id,
      signaling_type = "secreted",
      thresh = 0.01
    )
  })
)

sig_lr_secreted_all
```

# Contact and secreted significant pathways in all samples
```{r}
object.list.contact.single <- list(
  FVB = chat_fvb_contact,
  FVQ = chat_fvq_contact,
  FAM = chat_fam_contact,
  FCM = chat_fcm_contact,
  FEP = chat_fep_contact,
  FVS = chat_fvs_contact
)

object.list.secreted.single <- list(
  FVB = chat_fvb_secreted,
  FVQ = chat_fvq_secreted,
  FAM = chat_fam_secreted,
  FCM = chat_fcm_secreted,
  FEP = chat_fep_secreted,
  FVS = chat_fvs_secreted
)

# Empty plot if no signal
empty_rank_plot <- function(sample_id, signaling_type, reason) {
  ggplot() +
    annotate(
      "text",
      x = 0,
      y = 0,
      label = paste0(sample_id, "\n", signaling_type, "\n", reason),
      size = 4
    ) +
    theme_void()
}

# Safe rankNet wrapper
rankNet_safe <- function(
  chat,
  sample_id,
  signaling.type,
  title,
  slot.name = "netP"
) {
  
  annotation_use <- tryCatch(
    unique(chat@LR$LRsig$annotation),
    error = function(e) character(0)
  )
  
  if (!(signaling.type %in% annotation_use)) {
    return(
      empty_rank_plot(
        sample_id = sample_id,
        signaling_type = signaling.type,
        reason = "not in this object"
      )
    )
  }
  
  if (slot.name == "netP") {
    pathways_use <- tryCatch(chat@netP$pathways, error = function(e) character(0))
    
    if (is.null(pathways_use) || length(pathways_use) == 0) {
      return(
        empty_rank_plot(
          sample_id = sample_id,
          signaling_type = signaling.type,
          reason = "no pathway signal"
        )
      )
    }
  }
  
  p <- tryCatch(
    {
      rankNet(
        chat,
        mode = "single",
        slot.name = slot.name,
        measure = "weight",
        signaling.type = signaling.type,
        title = title
      )
    },
    error = function(e) {
      empty_rank_plot(
        sample_id = sample_id,
        signaling_type = signaling.type,
        reason = "rankNet failed"
      )
    }
  )
  
  return(p)
}

# Generate rankNet plots for each sample
rank_plots_by_sample <- lapply(names(object.list.contact.single), function(sample_id) {
  
  p_contact <- rankNet_safe(
    chat = object.list.contact.single[[sample_id]],
    sample_id = sample_id,
    signaling.type = "Cell-Cell Contact",
    title = paste0(sample_id, " Cell-Cell Contact"),
    slot.name = "netP"
  )
  
  p_secreted <- rankNet_safe(
    chat = object.list.secreted.single[[sample_id]],
    sample_id = sample_id,
    signaling.type = "Secreted Signaling",
    title = paste0(sample_id, " Secreted Signaling"),
    slot.name = "netP"
  )
  
  p_contact + p_secreted
})

names(rank_plots_by_sample) <- names(object.list.contact.single)

# Show each sample separately
rank_plots_by_sample$FVB
rank_plots_by_sample$FVQ
rank_plots_by_sample$FAM
rank_plots_by_sample$FCM
rank_plots_by_sample$FEP
rank_plots_by_sample$FVS
```


# Visualize the enriched cell groups
Contact
```{r}
sample.order <- c("FVB", "FVQ", "FAM", "FCM", "FEP", "FVS")

object.list.contact.single <- object.list.contact.single[sample.order]

plot_contact_hotspot <- function(chat, sample_id) {
  idx.Juxtracrine <- which(chat@LR$LRsig$annotation == "Cell-Cell Contact")
  
  chat <- spatialGiPlot(
    chat,
    plot.title = "Cell-Cell Contact",
    slot.name = "net",
    signaling.name = idx.Juxtracrine,
    measure = c("indeg"),
    measure.name = c("incoming"),
    show.plot = FALSE
  )
  
  chat@meta <- chat@meta %>%
    mutate(ContactHotspot = Getis.Ord.Gi.Cell.Cell.Contact * (Getis.Ord.Gi.P.Cell.Cell.Contact < 0.05)) %>%
    mutate(ContactHotspot = ifelse(ContactHotspot > 0, "Hot", ifelse(ContactHotspot < 0, "Cold", "Other")))
  
  tryCatch(
    spatialLeePlot(
      chat,
      group.by2 = "ContactHotspot",
      cutoff.Lee = 0.05,
      y.levels = c("Hot", "Cold", "Other")
    ),
    error = function(e) {
      message(sample_id, ": no finite ContactHotspot Lee matrix to plot")
      
      ggplot() +
        annotate(
          "text",
          x = 0,
          y = 0,
          label = paste0(sample_id, "\nNo finite ContactHotspot Lee matrix"),
          size = 4
        ) +
        theme_void()
    }
  )
}

contact_hotspot_plots <- lapply(names(object.list.contact.single), function(sample_id) {
  plot_contact_hotspot(
    object.list.contact.single[[sample_id]],
    sample_id
  )
})

names(contact_hotspot_plots) <- names(object.list.contact.single)

contact_hotspot_plots$FVB
contact_hotspot_plots$FVQ
contact_hotspot_plots$FAM
contact_hotspot_plots$FCM
contact_hotspot_plots$FEP
contact_hotspot_plots$FVS
```
Secreted
```{r}
sample.order <- c("FVB", "FVQ", "FAM", "FCM", "FEP", "FVS")

object.list.secreted.single <- object.list.secreted.single[sample.order]

plot_secreted_hotspot <- function(chat, sample_id) {
  idx.Paracrine <- which(chat@LR$LRsig$annotation == "Secreted Signaling")
  
  chat <- spatialGiPlot(
    chat,
    plot.title = "Secreted Signaling",
    slot.name = "net",
    signaling.name = idx.Paracrine,
    measure = c("indeg"),
    measure.name = c("incoming"),
    show.plot = FALSE
  )
  
  chat@meta <- chat@meta %>%
    mutate(SecretedHotspot = Getis.Ord.Gi.Secreted.Signaling * (Getis.Ord.Gi.P.Secreted.Signaling < 0.05)) %>%
    mutate(SecretedHotspot = ifelse(SecretedHotspot > 0, "Hot", ifelse(SecretedHotspot < 0, "Cold", "Other")))
  
  tryCatch(
    spatialLeePlot(
      chat,
      group.by2 = "SecretedHotspot",
      cutoff.Lee = 0.05,
      y.levels = c("Hot", "Cold", "Other")
    ),
    error = function(e) {
      message(sample_id, ": no finite SecretedHotspot Lee matrix to plot")
      
      ggplot() +
        annotate(
          "text",
          x = 0,
          y = 0,
          label = paste0(sample_id, "\nNo finite SecretedHotspot Lee matrix"),
          size = 4
        ) +
        theme_void()
    }
  )
}

secreted_hotspot_plots <- lapply(names(object.list.secreted.single), function(sample_id) {
  plot_secreted_hotspot(
    object.list.secreted.single[[sample_id]],
    sample_id
  )
})

names(secreted_hotspot_plots) <- names(object.list.secreted.single)

secreted_hotspot_plots$FVB
secreted_hotspot_plots$FVQ
secreted_hotspot_plots$FAM
secreted_hotspot_plots$FCM
secreted_hotspot_plots$FEP
secreted_hotspot_plots$FVS
```

# Heatmap of cell-type communication

```{r}
object.list.contact.single <- object.list.contact.single[sample.order]
object.list.secreted.single <- object.list.secreted.single[sample.order]

plot_celltype_heatmap <- function(chat, title.name, thresh = 0.05) {
  celltypes.use <- levels(chat@idents)
  
  comm <- tryCatch(
    subsetCommunication(
      chat,
      slot.name = "net",
      thresh = thresh
    ),
    error = function(e) {
      NULL
    }
  )
  
  if (is.null(comm) || nrow(comm) == 0) {
    return(
      ggplot() +
        annotate(
          "text",
          x = 0,
          y = 0,
          label = paste0(title.name, "\nNo significant interactions at threshold ", thresh),
          size = 4
        ) +
        theme_void()
    )
  }
  
  count.mat <- table(
    factor(comm$source, levels = celltypes.use),
    factor(comm$target, levels = celltypes.use)
  )
  
  weight.mat <- xtabs(
    prob ~ source + target,
    data = transform(
      comm,
      source = factor(source, levels = celltypes.use),
      target = factor(target, levels = celltypes.use)
    )
  )
  
  count.mat <- as.matrix(count.mat)
  weight.mat <- as.matrix(weight.mat)
  
  plot_one_heatmap <- function(mat, label, legend.title) {
    
    df <- as.data.frame(as.table(mat))
    colnames(df) <- c("source", "target", "value")
    
    df$source <- factor(df$source, levels = rev(rownames(mat)))
    df$target <- factor(df$target, levels = colnames(mat))
    
    if (legend.title == "Number") {
      df$label <- ifelse(df$value == 0, "", as.character(round(df$value)))
    } else {
      df$label <- ifelse(df$value == 0, "", as.character(signif(df$value, 2)))
    }
    
    values <- df$value[is.finite(df$value)]
    
    if (length(unique(values)) < 2) {
      return(
        ggplot() +
          annotate("text", x = 0, y = 0, label = paste0(title.name, "\n", label, "\nno variable values to plot"), size = 4) +
          theme_void()
      )
    }
    
    ggplot(df, aes(x = target, y = source, fill = value)) +
      geom_tile(color = "white", linewidth = 0.4) +
      geom_text(aes(label = label), size = 3) +
      scale_fill_gradient(low = "grey95", high = "#B2182B", name = legend.title) +
      labs(
        title = paste0(title.name, " - ", label),
        x = "Target",
        y = "Source"
      ) +
      coord_fixed() +
      theme_minimal(base_size = 12) +
      theme(
        plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
        axis.title = element_text(size = 11),
        axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 10),
        panel.grid = element_blank(),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9),
        plot.margin = margin(8, 8, 8, 8)
      )
  }
  
  plot_one_heatmap(count.mat, "significant interaction number", "Number") /
    plot_one_heatmap(weight.mat, "significant interaction strength", "Strength")
}
```

```{r, fig.width=6, fig.height=8}
plot_celltype_heatmap(chat_fvb_contact, "FVB contact")
plot_celltype_heatmap(chat_fvq_contact, "FVQ contact")
plot_celltype_heatmap(chat_fam_contact, "FAM contact")
plot_celltype_heatmap(chat_fcm_contact, "FCM contact")
plot_celltype_heatmap(chat_fep_contact, "FEP contact")
plot_celltype_heatmap(chat_fvs_contact, "FVS contact")
```

```{r, fig.width=6, fig.height=8}
plot_celltype_heatmap(chat_fvb_secreted, "FVB secreted")
plot_celltype_heatmap(chat_fvq_secreted, "FVQ secreted")
plot_celltype_heatmap(chat_fam_secreted, "FAM secreted")
plot_celltype_heatmap(chat_fcm_secreted, "FCM secreted")
plot_celltype_heatmap(chat_fep_secreted, "FEP secreted")
plot_celltype_heatmap(chat_fvs_secreted, "FVS secreted")
```
single cell pathway

# Single-cell pathway scores

```{r}
single_pathway_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/single_sample_cellchat/Single cell pathway"
sample.order <- c("FVB", "FVQ", "FAM", "FCM", "FEP", "FVS")

object.list.notch.single <- list(
  FVB = readRDS(file.path(single_pathway_dir, "chat_FVB_contactRange20_final.rds")),
  FVQ = readRDS(file.path(single_pathway_dir, "chat_FVQ_contactRange20_final.rds")),
  FAM = readRDS(file.path(single_pathway_dir, "chat_FAM_contactRange20_final.rds")),
  FCM = readRDS(file.path(single_pathway_dir, "chat_FCM_contactRange20_final.rds")),
  FEP = readRDS(file.path(single_pathway_dir, "chat_FEP_contactRange20_final.rds")),
  FVS = readRDS(file.path(single_pathway_dir, "chat_FVS_contactRange20_final.rds"))
)

object.list.tnf.single <- list(
  FVB = readRDS(file.path(single_pathway_dir, "chat_FVB_secreted_range250_final.rds")),
  FVQ = readRDS(file.path(single_pathway_dir, "chat_FVQ_secreted_range250_final.rds")),
  FAM = readRDS(file.path(single_pathway_dir, "chat_FAM_secreted_range250_final.rds")),
  FCM = readRDS(file.path(single_pathway_dir, "chat_FCM_secreted_range250_final.rds")),
  FEP = readRDS(file.path(single_pathway_dir, "chat_FEP_secreted_range250_final.rds")),
  FVS = readRDS(file.path(single_pathway_dir, "chat_FVS_secreted_range250_final.rds"))
)

object.list.notch.single <- object.list.notch.single[sample.order]
object.list.tnf.single <- object.list.tnf.single[sample.order]

region_celltype_cols <- c(
  "decidua eEVT" = "#B2182B",
  "decidua Endothelial" = "#2166AC",
  "decidua EVT" = "#1B7837",
  "muscle eEVT" = "#EF8A62",
  "muscle Endothelial" = "#67A9CF",
  "muscle EVT" = "#7FBF7B"
)

make_region_celltype_spatial_plot <- function(chat, sample_id, point.size = 1.2) {
  chat.region <- chat
  region.idents <- factor(
    paste(chat@meta$regions, as.character(chat@idents)),
    levels = names(region_celltype_cols)
  )
  names(region.idents) <- names(chat@idents)
  chat.region@idents <- region.idents
  
  SpatialCellChat::spatialDimPlot(
    chat.region,
    point.size = point.size
  ) +
    ggtitle(sample_id) +
    scale_color_manual(values = region_celltype_cols, limits = names(region_celltype_cols), drop = FALSE) +
    scale_fill_manual(values = region_celltype_cols, limits = names(region_celltype_cols), drop = FALSE)
}

compute_single_pathway_centrality <- function(chat) {
  SpatialCellChat::netAnalysis_computeCentrality(
    chat,
    slot.name = "netP",
    do.group = FALSE,
    degree.only = TRUE
  )
}

object.list.notch.single <- lapply(object.list.notch.single, compute_single_pathway_centrality)
object.list.tnf.single <- lapply(object.list.tnf.single, compute_single_pathway_centrality)

plot_single_pathway_score <- function(p_spatial, chat, pathway.show, sample_id = NULL, point.size = 1) {
  pathway.available <- if ("centr.cell" %in% names(chat@netP) &&
                           !is.null(dimnames(chat@netP$centr.cell)[[3]])) {
    dimnames(chat@netP$centr.cell)[[3]]
  } else if ("pathways.cell" %in% names(chat@netP)) {
    chat@netP$pathways.cell
  } else {
    chat@netP$pathways
  }
  
  if (!pathway.show %in% pathway.available) {
    message(pathway.show, " is not available for cell-level scoring in this CellChat object.")
    return(NULL)
  }
  
  if (is.null(sample_id)) {
    sample_id <- unique(chat@meta$donor)[1]
  }
  
  p_region_celltype <- make_region_celltype_spatial_plot(
    chat,
    sample_id = sample_id,
    point.size = point.size
  )
    
  g1 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    measure = c("outdeg"),
    do.binary = FALSE,
    color.heatmap = "Blues",
    point.size = point.size
  )
  
  g2 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    measure = c("indeg"),
    do.binary = FALSE,
    color.heatmap = "Reds",
    point.size = point.size
  )
  
  g3 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    merge = TRUE,
    do.binary = FALSE,
    color.heatmap = c("Blues", "Reds", "Purples"),
    point.size = point.size
  )
  
  patchwork::wrap_plots(
    p_region_celltype,
    g1,
    g2,
    g3,
    ncol = 4
  )
}
```

## NOTCH single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.notch.single$FVB, "NOTCH", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.notch.single$FVQ, "NOTCH", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.notch.single$FAM, "NOTCH", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.notch.single$FCM, "NOTCH", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.notch.single$FEP, "NOTCH", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.notch.single$FVS, "NOTCH", sample_id = "FVS")
```

## TNF single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.tnf.single$FVB, "TNF", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.tnf.single$FVQ, "TNF", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.tnf.single$FAM, "TNF", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.tnf.single$FCM, "TNF", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.tnf.single$FEP, "TNF", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.tnf.single$FVS, "TNF", sample_id = "FVS")
```
## IL1 single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.tnf.single$FVB, "IL1", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.tnf.single$FVQ, "IL1", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.tnf.single$FAM, "IL1", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.tnf.single$FCM, "IL1", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.tnf.single$FEP, "IL1", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.tnf.single$FVS, "IL1", sample_id = "FVS")
```

## VEGF single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.tnf.single$FVB, "VEGF", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.tnf.single$FVQ, "VEGF", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.tnf.single$FAM, "VEGF", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.tnf.single$FCM, "VEGF", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.tnf.single$FEP, "VEGF", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.tnf.single$FVS, "VEGF", sample_id = "FVS")
```

## NCAM single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.notch.single$FVB, "NCAM", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.notch.single$FVQ, "NCAM", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.notch.single$FAM, "NCAM", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.notch.single$FCM, "NCAM", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.notch.single$FEP, "NCAM", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.notch.single$FVS, "NCAM", sample_id = "FVS")
```

## PD-L1 single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.notch.single$FVB, "PD-L1", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.notch.single$FVQ, "PD-L1", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.notch.single$FAM, "PD-L1", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.notch.single$FCM, "PD-L1", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.notch.single$FEP, "PD-L1", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.notch.single$FVS, "PD-L1", sample_id = "FVS")
```

## LEP single-cell pathway scores

```{r, fig.width=16, fig.height=4}
plot_single_pathway_score(p_fvb, object.list.tnf.single$FVB, "LEP", sample_id = "FVB")
plot_single_pathway_score(p_fvq, object.list.tnf.single$FVQ, "LEP", sample_id = "FVQ")
plot_single_pathway_score(p_fam, object.list.tnf.single$FAM, "LEP", sample_id = "FAM")
plot_single_pathway_score(p_fcm, object.list.tnf.single$FCM, "LEP", sample_id = "FCM")
plot_single_pathway_score(p_fep, object.list.tnf.single$FEP, "LEP", sample_id = "FEP")
plot_single_pathway_score(p_fvs, object.list.tnf.single$FVS, "LEP", sample_id = "FVS")
```

## IL2 single-cell pathway scores NO

## IL4 single-cell pathway scores NO

