---
title: "Six Donor"
author:
- name: Zack Guo
  affiliation: UCL Institute of Child Health
  email: wegmzgu@ucl.ac.uk
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float: yes
    toc_depth: '5'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the library
```{r}
library(SpatialCellChat)
library(Matrix)
library(patchwork)
library(dplyr)
library(ggplot2)
library(Seurat)
options(stringsAsFactors = F)

# SpatialCellChat progressr workaround for R Markdown
future::plan("sequential")
```

#Read six sample RDS, where there are two control and six diseases
```{r}
obj_six <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_six_donors_Myometrium.rds"
)

# Check metadata columns
colnames(obj_six@meta.data)

# Check assay names
Assays(obj_six)

# Set Xenium as default assay
DefaultAssay(obj_six) <- "Xenium"

#Check number of cells from each donor
sample_col <- "donor"

table(obj_six@meta.data[[sample_col]])
```
#Function for all calculation in SpatialCellChat
```{r}
run_contact_cellchat_one_sample <- function(dataset, signaling.type) {
  
  ###Part I: Data input & processing and initialization of Spatial CellChat object
  
  
  ##Load the data
  
  #Load RDS
  obj_control_fvb <- dataset
  
  # Set Xenium as the active assay
  DefaultAssay(obj_control_fvb) <- "Xenium"
  
  # Generate normalized log-expression data
  obj_control_fvb <- NormalizeData(
    obj_control_fvb,
    assay = "Xenium",
    normalization.method = "LogNormalize",
    scale.factor = 10000,
    verbose = FALSE
  )
  
  #Prepare spatial coordinates of cells
  spatial.locs_control_fvb <- obj_control_fvb@meta.data[
    colnames(obj_control_fvb),
    c("centroid_X_1", "centroid_X_2"),
    drop = FALSE
  ]
  
  #Change name of columns to x and y
  colnames(spatial.locs_control_fvb) <- c("x", "y")
  
  #Define Xenium spatial factor. The spatial factor is already 1 so the conversion factor is just 1.
  conversion.factor <- 1
  spot.size <- 10
  
  #$tol=5 means it allows a 5 μm tolerance when defining spatial proximity.
  spatial.factors_control_fvb <- list(
    ratio = conversion.factor,
    tol = spot.size/2
  )
  
  
  ##Create a Spatial CellChat object
  
  chat_control_fvb <- createSpatialCellChat(
    object = obj_control_fvb,                  # Xenium Seurat object
    group.by = "CellTypeManual.l2",            # cell-type annotation
    assay = "Xenium",                          # normalized Xenium assay
    datatype = "spatial",
    coordinates = spatial.locs_control_fvb,    # cell centroid coordinates
    spatial.factors = spatial.factors_control_fvb
  )
  
  # Inspect the SpatialCellChat object
  chat_control_fvb
  
  # Check cell identities used in the analysis. x_cent/y_cent means x/y coordinate of cell centroid.
  table(chat_control_fvb@idents)
  
  
  ##Set the ligand-receptor interaction database
  
  custom_lr_clec11a_kit <- data.frame(
    ligand = "CLEC11A",
    receptor = "KIT",
    
    # Assign an individual pathway name so that the pair can be
    # retained in pathway-level CellChat analysis
    pathway_name = "CLEC11A",
    
    interaction_name = "CLEC11A_KIT",
    interaction_name_2 = "CLEC11A - KIT",
    
    # CLEC11A is treated as a soluble ligand
    annotation = "Secreted Signaling",
    
    evidence = paste(
      "NicheNet-predicted EVT-to-artery interaction;",
      "Greenbaum et al., Nature 2023, Fig. 6h;",
      "doi:10.1038/s41586-023-06298-9"
    ),
    
    stringsAsFactors = FALSE
  )
  
  # Merge the custom pair with the built-in human CellChat database
  CellChatDB_custom <- SpatialCellChat::updateCellChatDB(
    db = custom_lr_clec11a_kit,
    merged = TRUE,
    species_target = "human"
  )
  
  CellChatDB_use <- SpatialCellChat::subsetDB(
    CellChatDB_custom,
    search = c(
      "Secreted Signaling",
      "ECM-Receptor",
      "Cell-Cell Contact"
    ),
    non_protein = FALSE
  )
  
  
  ##Pre-processing the expression data for cell-cell communication analysis
  
  # Only keep contact signalling
  # Only keep genes in LR database.
  # Retain genes detected in at least 10 cells
  # Normally only spatially variable genes will be retain, but for Xenium data that has relatively small panel, we skip this
  
  
  # Select the human Cell-Cell Contact database
  CellChatDB <- CellChatDB.human
  
  CellChatDB_contact <- subsetDB(
    CellChatDB,
    search = "Cell-Cell Contact",
    non_protein = FALSE
  )
  
  CellChatDB_secreted <- subsetDB(
    CellChatDB_custom,
    search = "Secreted Signaling",
    non_protein = FALSE
  )
  
  CellChatDB_selected <- switch(
    signaling.type,
    contact = CellChatDB_contact,
    secreted = CellChatDB_secreted
  )
  
  # Assign the selected database to the SpatialCellChat object
  chat_control_fvb@DB <- CellChatDB_selected
  
  
  # Subset expression data to signalling genes
  # This step is required even when using the whole database
  chat_control_fvb <- subsetData(chat_control_fvb)
  
  
  # Count how many cells express each signalling gene
  # Do this BEFORE preProcessing/ALRA imputation
  n_cells_expressed <- Matrix::rowSums(
    chat_control_fvb@data.signaling > 0
  )
  
  
  # Retain genes detected in at least 10 cells
  min_cells_gene <- 10
  
  features_small_panel <- names(
    n_cells_expressed[n_cells_expressed >= min_cells_gene]
  )
  
  
  # Run SpatialCellChat preprocessing
  # This performs ALRA preprocessing on the signalling-expression matrix to make it compatible for spatial cellchat function
  chat_control_fvb <- preProcessing(chat_control_fvb)
  
  
  # Manually assign the genes retained for the small Xenium panel
  # This replaces identifyOverExpressedGenes()
  chat_control_fvb@var.features[["features"]] <- features_small_panel
  
  chat_control_fvb@var.features[["features.info"]] <- data.frame(
    features = names(n_cells_expressed),
    n_cells_expressed = as.numeric(n_cells_expressed),
    selected = n_cells_expressed >= min_cells_gene,
    row.names = names(n_cells_expressed),
    stringsAsFactors = FALSE
  )
  
  
  # Identify available ligand-receptor interactions
  # variable.both = FALSE means either the ligand or receptor must pass
  # the >=10-cell filter, while all required genes must be in the panel
  chat_control_fvb <- identifyOverExpressedInteractions(
    chat_control_fvb,
    variable.both = FALSE
  )
  
  
  ###Part II: Inference of cell-cell communication network
  
  
  ##Compute the communication probability and infer cellular communication network at the individual cell level(L-R)
  
  # Check the retained ligand-receptor pairs
  nrow(chat_control_fvb@LR$LRsig)
  
  # Confirm that the current analysis contains only contact signalling
  unique(chat_control_fvb@LR$LRsig$annotation)
  
  # Compute individual-cell LR communication probabilities using spatial distance
  chat_control_fvb <- switch(
    signaling.type,
    
    contact = computeCommunProb(
      chat_control_fvb,
      distance.use = TRUE,
      scale.distance = 0.8,
      contact.dependent = TRUE,
      interaction.range = 250,
      contact.range = 20
    ),
    
    secreted = computeCommunProb(
      chat_control_fvb,
      distance.use = TRUE,
      scale.distance = 0.8,
      contact.dependent = FALSE,
      interaction.range = 250
    )
  )
  
  
  # Filtering
  
  # Filter weak or poorly supported individual-cell LR communication links
  chat_control_fvb <- filterProbability(
    chat_control_fvb
  )
  
  # Keep only communication events with sufficient cell-cell link support
  chat_control_fvb <- filterCommunication(
    chat_control_fvb,
    min.cells = NULL,
    min.links = 10,
    min.cells.sr = 10
  )
  
  
  ## Calculate group-level LR communication probability
  
  # Average filtered individual-cell LR probabilities into source-target cell-group LR probabilities
  chat_control_fvb <- computeAvgCommunProb(
    chat_control_fvb,
    avg.type = "avg",
    min.percent = 0.1,
    min.cells.sr = 10,
    do.permutation = TRUE,
    nboot = 100,
    seed.use = 1L
  )
  
  # Remove group-level LR communications involving cell groups with fewer than 10 cells
  chat_control_fvb <- filterCommunication(
    chat_control_fvb,
    min.cells = 10,
    min.links = NULL,
    min.cells.sr = NULL
  )
  
  
  ##Infer communication at signaling-pathway level at cell group level
  
  chat_control_fvb <- computeCommunProbPathway(
    chat_control_fvb,
    thresh = 0.05,
    do.group = TRUE,
    do.cell = FALSE
  )
  
  
  ##Calculate the aggregated communication network
  
  # Aggregate LR/pathway communication probabilities into cell-group-level count and weight networks
  chat_control_fvb <- aggregateNet(
    chat_control_fvb
  )
  
  
  ##Compute network centrality scores
  
  # Signalling role of each cell type. Sender? Receiver?
  
  chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
    chat_control_fvb,
    slot.name = "net",
    do.group = FALSE,
    degree.only = TRUE
  )
  
  chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
    chat_control_fvb,
    slot.name = "net",
    do.group = TRUE,
    degree.only = TRUE
  )
  
  try(
    chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
      chat_control_fvb,
      slot.name = "netP",
      do.group = TRUE,
      degree.only = TRUE
    ),
    silent = TRUE
  )
  
  
  ##Return final SpatialCellChat object
  
  return(chat_control_fvb)
}
```

#Function for saving RDS
```{r}
save_cellchat_one_sample <- function(chat_object, sample_id, signaling.type) {
  
  output_data_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/single_sample_cellchat"
  
  dir.create(
    output_data_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
  
  sample_clean <- gsub("[^A-Za-z0-9_\\-]", "_", sample_id)
  
  file_suffix <- switch(
    signaling.type,
    contact = "contactRange20",
    secreted = "secreted_range250"
  )
  
  output_file <- file.path(
    output_data_dir,
    paste0(
      "chat_",
      sample_clean,
      "_",
      file_suffix,
      "_final.rds"
    )
  )
  
  saveRDS(
    chat_object,
    output_file
  )
  
  return(output_file)
}
```

#Run for FCM sample
```{r}
obj_fcm <- subset(
  obj_six,
  subset = donor == "FCM"
)

chat_fcm_contact <- run_contact_cellchat_one_sample(
  obj_fcm,
  signaling.type = "contact"
)

chat_fcm_secreted <- run_contact_cellchat_one_sample(
  obj_fcm,
  signaling.type = "secreted"
)
```

# Save FCM contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fcm_contact,
  sample_id = "FCM",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fcm_secreted,
  sample_id = "FCM",
  signaling.type = "secreted"
)
```

# Run for FEP sample
```{r}
obj_fep <- subset(
  obj_six,
  subset = donor == "FEP"
)

chat_fep_contact <- run_contact_cellchat_one_sample(
  obj_fep,
  signaling.type = "contact"
)

chat_fep_secreted <- run_contact_cellchat_one_sample(
  obj_fep,
  signaling.type = "secreted"
)
```

# Save FEP contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fep_contact,
  sample_id = "FEP",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fep_secreted,
  sample_id = "FEP",
  signaling.type = "secreted"
)
```

# Run for FVQ sample
```{r}
obj_fvq <- subset(
  obj_six,
  subset = donor == "FVQ"
)

chat_fvq_contact <- run_contact_cellchat_one_sample(
  obj_fvq,
  signaling.type = "contact"
)

chat_fvq_secreted <- run_contact_cellchat_one_sample(
  obj_fvq,
  signaling.type = "secreted"
)
```

# Save FVQ contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fvq_contact,
  sample_id = "FVQ",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvq_secreted,
  sample_id = "FVQ",
  signaling.type = "secreted"
)
```

# Run for FVS sample
```{r}
obj_fvs <- subset(
  obj_six,
  subset = donor == "FVS"
)

chat_fvs_contact <- run_contact_cellchat_one_sample(
  obj_fvs,
  signaling.type = "contact"
)

chat_fvs_secreted <- run_contact_cellchat_one_sample(
  obj_fvs,
  signaling.type = "secreted"
)
```

# Save FVS contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fvs_contact,
  sample_id = "FVS",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvs_secreted,
  sample_id = "FVS",
  signaling.type = "secreted"
)
```
# Run for FVB sample
```{r}
obj_fvb <- subset(
  obj_six,
  subset = donor == "FVB"
)

chat_fvb_contact <- run_contact_cellchat_one_sample(
  obj_fvb,
  signaling.type = "contact"
)

chat_fvb_secreted <- run_contact_cellchat_one_sample(
  obj_fvb,
  signaling.type = "secreted"
)
```

# Save FVB contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fvb_contact,
  sample_id = "FVB",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fvb_secreted,
  sample_id = "FVB",
  signaling.type = "secreted"
)
```

# Run for FAM sample
```{r}
obj_fam <- subset(
  obj_six,
  subset = donor == "FAM"
)

chat_fam_contact <- run_contact_cellchat_one_sample(
  obj_fam,
  signaling.type = "contact"
)

chat_fam_secreted <- run_contact_cellchat_one_sample(
  obj_fam,
  signaling.type = "secreted"
)
```

# Save FAM contact and secreted objects
```{r}
save_cellchat_one_sample(
  chat_object = chat_fam_contact,
  sample_id = "FAM",
  signaling.type = "contact"
)

save_cellchat_one_sample(
  chat_object = chat_fam_secreted,
  sample_id = "FAM",
  signaling.type = "secreted"
)
```

