---
title: "SpatialCellchat"
author: "Zack"
date: "2026-06-12"
output: html_document
knit: >-
  (function(inputFile, encoding) {
    progressr::handlers(global = TRUE)
    rmarkdown::render(
      input = inputFile,
      encoding = encoding,
      envir = new.env(parent = globalenv())
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the library
```{r}
library(SpatialCellChat)
library(Matrix)
library(patchwork)
library(dplyr)
library(ggplot2)
library(Seurat)
options(stringsAsFactors = F)

# SpatialCellChat progressr workaround for R Markdown
future::plan("sequential")
```

Part I: Data input & processing and initialization of Spatial CellChat object
Load the data
```{r,eval=FALSE}
#Load RDS
obj_control_fvb <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_donor_FVB_Myometrium.rds"
)

# Set Xenium as the active assay
DefaultAssay(obj_control_fvb) <- "Xenium"

# Generate normalized log-expression data
obj_control_fvb <- NormalizeData(
  obj_control_fvb,
  assay = "Xenium",
  normalization.method = "LogNormalize",
  scale.factor = 10000,
  verbose = FALSE
)

#Prepare spatial coordinates of cells
spatial.locs_control_fvb <- obj_control_fvb@meta.data[
  colnames(obj_control_fvb),
  c("centroid_X_1", "centroid_X_2"),
  drop = FALSE
]

#Change name of columns to x and y
colnames(spatial.locs_control_fvb) <- c("x", "y")

#Define Xenium spatial factor. The spatial factor is already 1 so the conversion factor is just 1.
conversion.factor <- 1
spot.size <- 10

#$tol=5 means it allows a 5 μm tolerance when defining spatial proximity.
spatial.factors_control_fvb <- list(
  ratio = conversion.factor,
  tol = spot.size/2
)
```

Create a Spatial CellChat object
```{r,eval=FALSE}
chat_control_fvb <- createSpatialCellChat(
  object = obj_control_fvb,                  # Xenium Seurat object
  group.by = "CellTypeManual.l2",            # cell-type annotation
  assay = "Xenium",                          # normalized Xenium assay
  datatype = "spatial",
  coordinates = spatial.locs_control_fvb,    # cell centroid coordinates
  spatial.factors = spatial.factors_control_fvb
)

# Inspect the SpatialCellChat object
chat_control_fvb

# Check cell identities used in the analysis. x_cent/y_cent means x/y coordinate of cell centroid.
table(chat_control_fvb@idents)
```

Set the ligand-receptor interaction database
```{r,eval=FALSE}
custom_lr_clec11a_kit <- data.frame(
  ligand = "CLEC11A",
  receptor = "KIT",
  
  # Assign an individual pathway name so that the pair can be
  # retained in pathway-level CellChat analysis
  pathway_name = "CLEC11A",
  
  interaction_name = "CLEC11A_KIT",
  interaction_name_2 = "CLEC11A - KIT",
  
  # CLEC11A is treated as a soluble ligand
  annotation = "Secreted Signaling",
  
  evidence = paste(
    "NicheNet-predicted EVT-to-artery interaction;",
    "Greenbaum et al., Nature 2023, Fig. 6h;",
    "doi:10.1038/s41586-023-06298-9"
  ),
  
  stringsAsFactors = FALSE
)

# Merge the custom pair with the built-in human CellChat database
CellChatDB_custom <- SpatialCellChat::updateCellChatDB(
  db = custom_lr_clec11a_kit,
  merged = TRUE,
  species_target = "human"
)

CellChatDB_use <- SpatialCellChat::subsetDB(
  CellChatDB_custom,
  search = c(
    "Secreted Signaling",
    "ECM-Receptor",
    "Cell-Cell Contact"
  ),
  non_protein = FALSE
)
```
Pre-processing the expression data for cell-cell communication analysis
```{r,eval=FALSE}
# 1. Select the human Cell-Cell Contact database
CellChatDB <- CellChatDB.human

CellChatDB_contact <- subsetDB(
  CellChatDB,
  search = "Cell-Cell Contact",
  non_protein = FALSE
)

# Assign the selected database to the SpatialCellChat object
chat_control_fvb@DB <- CellChatDB_contact


# 2. Subset expression data to signalling genes
# This step is required even when using the whole database
chat_control_fvb <- subsetData(chat_control_fvb)


# 3. Count how many cells express each signalling gene
# Do this BEFORE preProcessing/ALRA imputation
n_cells_expressed <- Matrix::rowSums(
  chat_control_fvb@data.signaling > 0
)


# 4. Retain genes detected in at least 10 cells
min_cells_gene <- 10

features_small_panel <- names(
  n_cells_expressed[n_cells_expressed >= min_cells_gene]
)


# 5. Run SpatialCellChat preprocessing
# This performs ALRA preprocessing on the signalling-expression matrix
chat_control_fvb <- preProcessing(chat_control_fvb)


# 6. Manually assign the genes retained for the small Xenium panel
# This replaces identifyOverExpressedGenes()
chat_control_fvb@var.features[["features"]] <- features_small_panel

chat_control_fvb@var.features[["features.info"]] <- data.frame(
  features = names(n_cells_expressed),
  n_cells_expressed = as.numeric(n_cells_expressed),
  selected = n_cells_expressed >= min_cells_gene,
  row.names = names(n_cells_expressed),
  stringsAsFactors = FALSE
)


# 7. Identify available ligand-receptor interactions
# variable.both = FALSE means either the ligand or receptor must pass
# the >=10-cell filter, while all required genes must be in the panel
chat_control_fvb <- identifyOverExpressedInteractions(
  chat_control_fvb,
  variable.both = FALSE
)
```
Part II: Inference of cell-cell communication network
Compute the communication probability and infer cellular communication network at the individual cell level
```{r expensive-computation, eval=FALSE}
# Check the retained ligand-receptor pairs
nrow(chat_control_fvb@LR$LRsig)

# Confirm that the current analysis contains only contact signalling
unique(chat_control_fvb@LR$LRsig$annotation)

# 1. Compute communication probabilities between individual cells
chat_control_fvb <- computeCommunProb(
  chat_control_fvb,
  distance.use = TRUE,
  scale.distance = 0.8,
  contact.dependent = TRUE,
  interaction.range = 250,
  contact.range = 20
)

#Checkpoint
output_data_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data"

dir.create(
  output_data_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_rds(
  chat_control_fvb,
  file.path(
    output_data_dir,
    "chat_control_fvb_contactRange20_unfiltered.rds"
  )
)
```

```{r,eval=FALSE}
#Load the checkpoint
chat_control_fvb <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_control_fvb_contactRange20_unfiltered.rds"
)

chat_control_fvb <- filterProbability(
  chat_control_fvb
)

chat_control_fvb <- filterCommunication(
  chat_control_fvb,
  min.cells = NULL,
  min.links = 10,
  min.cells.sr = 10
)
```

```{r,eval=FALSE}
# Load the checkpoint generated after computeCommunProb()
chat_control_fvb <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_control_fvb_contactRange20_unfiltered.rds"
)

# Filter the individual-cell LR communication network
chat_control_fvb <- filterProbability(
  chat_control_fvb
)

chat_control_fvb <- filterCommunication(
  chat_control_fvb,
  min.cells = NULL,
  min.links = 10,
  min.cells.sr = 10
)
```

Calculate group-level LR communication probability
```{r,eval=FALSE}
chat_control_fvb <- computeAvgCommunProb(
  chat_control_fvb,
  avg.type = "avg",
  min.percent = 0.1,
  min.cells.sr = 10,
  do.permutation = TRUE,
  nboot = 100,
  seed.use = 1L
)

chat_control_fvb <- filterCommunication(
  chat_control_fvb,
  min.cells = 10,
  min.links = NULL,
  min.cells.sr = NULL
)
```

Infer communication at signaling-pathway level
```{r,eval=FALSE}
chat_control_fvb <- computeCommunProbPathway(
  chat_control_fvb,
  thresh = 0.05,
  do.group = TRUE,
  do.cell = FALSE
)
```

Calculate the aggregated communication network
```{r,eval=FALSE}
chat_control_fvb <- aggregateNet(
  chat_control_fvb
)
```

Compute network centrality scores
```{r,eval=FALSE}
chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
  chat_control_fvb,
  slot.name = "net",
  do.group = FALSE,
  degree.only = TRUE
)

chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
  chat_control_fvb,
  slot.name = "net",
  do.group = TRUE,
  degree.only = TRUE
)

chat_control_fvb <- SpatialCellChat::netAnalysis_computeCentrality(
  chat_control_fvb,
  slot.name = "netP",
  do.group = TRUE,
  degree.only = TRUE
)
```

Load 4 processed data: Control contact. control secreted, dissease contact, disease secreted
```{r}
chat_disease_fam <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_disease_fam_contactRange20_final.rds"
)

chat_disease_fam_secreted <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_disease_fam_secreted_range250_final.rds"
)

chat_control_fvb <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_control_fvb_contactRange20_final.rds"
)

chat_control_fvb_secreted <- readr::read_rds(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/chat_control_fvb_secreted_range250_final.rds"
)
```

```{r}
sig_lr_control_contact <- SpatialCellChat::subsetCommunication(
  chat_control_fvb,
  slot.name = "net",
  thresh = 0.01
) %>%
  arrange(pval, desc(prob))

sig_lr_disease_contact <- SpatialCellChat::subsetCommunication(
  chat_disease_fam,
  slot.name = "net",
  thresh = 0.01
) %>%
  arrange(pval, desc(prob))

sig_lr_control_secreted <- SpatialCellChat::subsetCommunication(
  chat_control_fvb_secreted,
  slot.name = "net",
  thresh = 0.01
) %>%
  arrange(pval, desc(prob))

sig_lr_disease_secreted <- SpatialCellChat::subsetCommunication(
  chat_disease_fam_secreted,
  slot.name = "net",
  thresh = 0.01
) %>%
  arrange(pval, desc(prob))
```

```{r}
sig_lr_control_contact
sig_lr_disease_contact

sig_lr_control_secreted
sig_lr_disease_secreted
```


identity check
```{r}
# Load original Seurat objects
obj_control_fvb <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_donor_FVB_Myometrium.rds"
)

obj_disease_fam <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_donor_FAM_Myometrium.rds"
)
```

```{r}
marker_panels <- list(
  `immune genes` = c(
    "CTLA4", "PDCD1",
    "NKG7", "KLRD1", "GNLY",
    "LST1", "TYROBP", "FCER1G",
    "LILRB2", "MRC1",
    "CEACAM8", "CCR3",
    "CCR4", "CCR5", "CCR7",
    "CXCR5", "CXCR6"
  ),

  `chemokine program` = c(
    "TNF",
    "CCL3", "CCL3L1",
    "CCL4", "CCL5",
    "CXCL13"
  ),

  `MHC-I / lineage check` = c(
    "HLA-G", "HLA-B",
    "CD8A", "CD8B",
    "PTPRC", "TRAC", "CD3D", "CD3E",
    "KRT7", "ITGA5", "MMP2",
    "PECAM1", "VWF"
  )
)

```

```{r}
make_identity_check <- function(obj, plot_title) {

  DefaultAssay(obj) <- "Xenium"

  obj_sub <- subset(
    obj,
    subset = CellTypeManual.l2 %in% c("eEVT", "Endothelial")
  )

  obj_sub$CellTypeManual.l2 <- factor(
    obj_sub$CellTypeManual.l2,
    levels = c("eEVT", "Endothelial")
  )

  panel_genes <- rownames(obj_sub[["Xenium"]])

  markers_present <- lapply(
    marker_panels,
    function(x) intersect(x, panel_genes)
  )

  # Remove empty marker categories
  markers_present <- markers_present[lengths(markers_present) > 0]

  DotPlot(
    obj_sub,
    features = markers_present,
    group.by = "CellTypeManual.l2",
    assay = "Xenium",
    scale = TRUE,
    dot.scale = 6
  ) +
    RotatedAxis() +
    labs(
      title = plot_title,
      x = "Marker genes",
      y = "Cell identity"
    ) +
    theme(
      axis.text.x = element_text(
        angle = 45,
        hjust = 1,
        size = 8
      ),
      axis.text.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5),
      strip.text = element_text(size = 8)
    )
}
```

```{r}
p_control_identity <- make_identity_check(
  obj_control_fvb,
  "Control FVB: eEVT and Endothelial identity check"
)

p_control_identity

p_disease_identity <- make_identity_check(
  obj_disease_fam,
  "Disease FAM: eEVT and Endothelial identity check"
)

p_disease_identity
```
```{r}
key_gene_panels <- list(
  `Contact LR hits` = c(
    "JAG1", "NOTCH2", "NOTCH3",
    "CD274", "PDCD1LG2", "PDCD1",   # PD-L1, PD-L2, PD-1
    "CEACAM1",
    "CDH5"
  ),
  
  `Secreted LR hits` = c(
    "CCL5", "CCL21", "CCL28",
    "CCR3", "CCR7",
    "SPP1", "CD44",
    "CSF1", "CSF1R",
    
    # IL1 / IL2 / IL4 signaling
    "IL1B", "IL1R1", "IL1RAP",
    "IL2", "IL2RG",
    "IL4", "IL4R"
  )
)

make_key_gene_check <- function(obj, plot_title) {
  
  DefaultAssay(obj) <- "Xenium"
  
  obj_sub <- subset(
    obj,
    subset = CellTypeManual.l2 %in% c("eEVT", "Endothelial")
  )
  
  obj_sub$CellTypeManual.l2 <- factor(
    obj_sub$CellTypeManual.l2,
    levels = c("eEVT", "Endothelial")
  )
  
  panel_genes <- rownames(obj_sub[["Xenium"]])
  
  key_genes_present <- lapply(
    key_gene_panels,
    function(x) intersect(x, panel_genes)
  )
  
  key_genes_present <- key_genes_present[lengths(key_genes_present) > 0]
  
  DotPlot(
    obj_sub,
    features = key_genes_present,
    group.by = "CellTypeManual.l2",
    assay = "Xenium",
    scale = TRUE,
    dot.scale = 3.2,     # smaller dots
    dot.min = 0.01       # remove extremely tiny background dots
  ) +
    RotatedAxis() +
    labs(
      title = plot_title,
      x = "Key genes",
      y = "Cell type"
    ) +
    theme(
      axis.text.x = element_text(
        angle = 45,
        hjust = 1,
        vjust = 1,
        size = 7
      ),
      axis.text.y = element_text(size = 9),
      axis.title.x = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 11),
      strip.text = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.title = element_text(size = 8),
      panel.spacing = unit(0.8, "lines")
    )
}

p_control_key <- make_key_gene_check(
  obj_control_fvb,
  "Control FVB: key LR gene check"
)

p_disease_key <- make_key_gene_check(
  obj_disease_fam,
  "Disease FAM: key LR gene check"
)

p_control_key
p_disease_key
```

```{r}
spatialCCCDistPlot(chat_control_fvb)
```


