---
title: "CellChat analysis of multiple spatial transcriptomics datasets"
author:
- name: Zack Guo
  affiliation: UCL Institute of Child Health
  email: wegmzgu@ucl.ac.uk
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float: yes
    toc_depth: '5'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the library
```{r}
library(SpatialCellChat)
library(CellChat)
library(Matrix)
library(patchwork)
library(dplyr)
library(ggplot2)
library(Seurat)
options(stringsAsFactors = F)

# SpatialCellChat progressr workaround for R Markdown
future::plan("sequential")
```
#Load data
```{r}
obj_six <- readRDS(
  "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Data/Spatial_Xenium_six_donors_Myometrium.rds"
)
```

#Main function for processing and calculating interaction for multiple spatial dataset
```{r}
run_multisample_condition_cellchat <- function(dataset, donors_use, signaling.type) {
  
  ###Part I: Data input & processing and initialization of CellChat object
  
  
  ##Load data
  
  obj_condition <- subset(
    dataset,
    subset = donor %in% donors_use
  )
  
  # Set Xenium as the active assay
  DefaultAssay(obj_condition) <- "Xenium"
  
  # Generate normalized log-expression data
  obj_condition <- NormalizeData(
    obj_condition,
    assay = "Xenium",
    normalization.method = "LogNormalize",
    scale.factor = 10000,
    verbose = FALSE
  )
  
  # Join Seurat v5 layers into a single assay layer
  obj_condition <- JoinLayers(
    obj_condition,
    assay = "Xenium"
  )
  
  
  ##Prepare input data for CellChat analysis
  
  # a column named `samples` should be provided for spatial transcriptomics analysis,
  # which is useful for analyzing cell-cell communication by aggregating multiple samples/replicates.
  obj_condition$samples <- factor(
    obj_condition$donor,
    levels = donors_use
  )
  
  # define the meta data
  obj_condition$labels <- factor(
    obj_condition$CellTypeManual.l2
  )
  
  # Check the cell labels
  unique(obj_condition$labels)
  
  # Check the sample labels
  unique(obj_condition$samples)
  
  
  ##Load spatial transcriptomics information
  
  # Spatial locations of cells are required
  spatial.locs <- obj_condition@meta.data[
    colnames(obj_condition),
    c("centroid_X_1", "centroid_X_2"),
    drop = FALSE
  ]
  
  colnames(spatial.locs) <- c("x", "y")
  
  # Scale factors of spatial coordinates
  conversion.factor <- 1
  spot.size <- 10
  
  spatial.factors <- data.frame(
    ratio = rep(conversion.factor, length(donors_use)),
    tol = rep(spot.size/2, length(donors_use))
  )
  
  rownames(spatial.factors) <- donors_use
  
  
  ##Create a CellChat object
  
  cellchat <- SpatialCellChat::createSpatialCellChat(
    object = obj_condition,
    group.by = "labels",
    assay = "Xenium",
    datatype = "spatial",
    coordinates = spatial.locs,
    spatial.factors = spatial.factors
  )
  
  cellchat
  
  
  ##Set the ligand-receptor interaction database
  
  custom_lr_clec11a_kit <- data.frame(
    ligand = "CLEC11A",
    receptor = "KIT",
    
    # Assign an individual pathway name so that the pair can be
    # retained in pathway-level CellChat analysis
    pathway_name = "CLEC11A",
    
    interaction_name = "CLEC11A_KIT",
    interaction_name_2 = "CLEC11A - KIT",
    
    # CLEC11A is treated as a soluble ligand
    annotation = "Secreted Signaling",
    
    evidence = paste(
      "NicheNet-predicted EVT-to-artery interaction;",
      "Greenbaum et al., Nature 2023, Fig. 6h;",
      "doi:10.1038/s41586-023-06298-9"
    ),
    
    stringsAsFactors = FALSE
  )
  
  # Merge the custom pair with the built-in human CellChat database
  CellChatDB_custom <- SpatialCellChat::updateCellChatDB(
    db = custom_lr_clec11a_kit,
    merged = TRUE,
    species_target = "human"
  )
  
  CellChatDB <- CellChatDB.human # use CellChatDB.human if running on human data
  
  # use a subset of CellChatDB for cell-cell communication analysis
  CellChatDB_contact <- SpatialCellChat::subsetDB(
    CellChatDB,
    search = "Cell-Cell Contact",
    key = "annotation"
  )
  
  CellChatDB_secreted <- SpatialCellChat::subsetDB(
    CellChatDB_custom,
    search = "Secreted Signaling",
    key = "annotation"
  )
  
  CellChatDB_use <- switch(
    signaling.type,
    contact = CellChatDB_contact,
    secreted = CellChatDB_secreted
  )
  
  # set the used database in the object
  cellchat@DB <- CellChatDB_use
  
  
  ##Preprocessing the expression data for cell-cell communication analysis
  
  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
  
  # Count how many cells express each signalling gene
  # Do this BEFORE preProcessing/ALRA imputation
  n_cells_expressed <- Matrix::rowSums(
    cellchat@data.signaling > 0
  )
  
  # Retain genes detected in at least 10 cells
  min_cells_gene <- 10
  
  features_small_panel <- names(
    n_cells_expressed[n_cells_expressed >= min_cells_gene]
  )
  
  # Run SpatialCellChat preprocessing
  cellchat <- preProcessing(cellchat)
  
  # Manually assign the genes retained for the small Xenium panel
  # This replaces identifyOverExpressedGenes()
  cellchat@var.features[["features"]] <- features_small_panel
  
  cellchat@var.features[["features.info"]] <- data.frame(
    features = names(n_cells_expressed),
    n_cells_expressed = as.numeric(n_cells_expressed),
    selected = n_cells_expressed >= min_cells_gene,
    row.names = names(n_cells_expressed),
    stringsAsFactors = FALSE
  )
  
  # identify over-expressed ligand-receptor interactions
  cellchat <- identifyOverExpressedInteractions(cellchat)
  
  nrow(cellchat@LR$LRsig)
  unique(cellchat@LR$LRsig$annotation)
  
  
  ###Part II: Inference of cell-cell communication network
  
  
  ##Compute the communication probability and infer cellular communication network
  
  cellchat <- switch(
    signaling.type,
    
    contact = SpatialCellChat::computeCommunProb(
      cellchat,
      distance.use = TRUE,
      scale.distance = 3.4,
      contact.dependent = TRUE,
      interaction.range = 250,
      contact.range = 20
    ),
    
    secreted = SpatialCellChat::computeCommunProb(
      cellchat,
      distance.use = TRUE,
      scale.distance = 3.4,
      contact.dependent = FALSE,
      interaction.range = 250
    )
  )
  
  
  ##Filtering

  cellchat <- filterProbability(cellchat)
  
  cellchat <- filterCommunication(
    cellchat,
    min.cells = NULL,
    min.links = 10,
    min.cells.sr = 10
  )
  
  
  ##Calculate group-level LR communication probability
  
  cellchat <- computeAvgCommunProb(
    cellchat,
    avg.type = "avg",
    min.percent = 0.1,
    min.cells.sr = 10,
    do.permutation = TRUE,
    nboot = 100,
    seed.use = 1L
  )
  
  cellchat <- filterCommunication(
    cellchat,
    min.cells = 10,
    min.links = NULL,
    min.cells.sr = NULL
  )
  
  
  ##Infer the cell-cell communication at a signaling pathway level
  
  cellchat <- computeCommunProbPathway(
    cellchat,
    thresh = 0.05,
    do.group = TRUE,
    do.cell = FALSE
  )
  
  
  ##Calculate the aggregated cell-cell communication network
  
  cellchat <- aggregateNet(cellchat)
  
  
  ##Compute network centrality scores
  
  cellchat <- SpatialCellChat::netAnalysis_computeCentrality(
    cellchat,
    slot.name = "net",
    do.group = TRUE,
    degree.only = TRUE
  )
  
  try(
    cellchat <- SpatialCellChat::netAnalysis_computeCentrality(
      cellchat,
      slot.name = "netP",
      do.group = TRUE,
      degree.only = TRUE
    ),
    silent = TRUE
  )
  
  
  ##Return final CellChat object
  
  return(cellchat)
}
```

#Apply function to control group
```{r}
cellchat.control.contact <- run_multisample_condition_cellchat(
  dataset = obj_six,
  donors_use = c("FVB", "FVQ"),
  signaling.type = "contact"
)

cellchat.control.secreted <- run_multisample_condition_cellchat(
  dataset = obj_six,
  donors_use = c("FVB", "FVQ"),
  signaling.type = "secreted"
)

cellchat.disease.contact <- run_multisample_condition_cellchat(
  dataset = obj_six,
  donors_use = c("FCM", "FEP", "FVS", "FAM"),
  signaling.type = "contact"
)

cellchat.disease.secreted <- run_multisample_condition_cellchat(
  dataset = obj_six,
  donors_use = c("FCM", "FEP", "FVS", "FAM"),
  signaling.type = "secreted"
)
```
#Save RDS
```{r}
output_data_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat"

saveRDS(
  cellchat.control.contact,
  file.path(
    output_data_dir,
    "cellchat_control_multi_contactRange20_final.rds"
  )
)

saveRDS(
  cellchat.control.secreted,
  file.path(
    output_data_dir,
    "cellchat_control_multi_secreted_range250_final.rds"
  )
)

saveRDS(
  cellchat.disease.contact,
  file.path(
    output_data_dir,
    "cellchat_disease_multi_contactRange20_final.rds"
  )
)

saveRDS(
  cellchat.disease.secreted,
  file.path(
    output_data_dir,
    "cellchat_disease_multi_secreted_range250_final.rds"
  )
)
```

# Visualize multi-sample pathway scores

```{r}
multi_sample_dir <- "C:/Bioinfo General/Spatial Analysis/Spatial analysis project/Output/Data/multi_sample_cellchat"

cellchat.control.contact <- readRDS(
  file.path(multi_sample_dir, "cellchat_control_multi_contactRange20_final.rds")
)

cellchat.disease.contact <- readRDS(
  file.path(multi_sample_dir, "cellchat_disease_multi_contactRange20_final.rds")
)

cellchat.control.secreted <- readRDS(
  file.path(multi_sample_dir, "cellchat_control_multi_secreted_range250_final.rds")
)

cellchat.disease.secreted <- readRDS(
  file.path(multi_sample_dir, "cellchat_disease_multi_secreted_range250_final.rds")
)

compute_multi_single_pathway_centrality <- function(chat) {
  SpatialCellChat::netAnalysis_computeCentrality(
    chat,
    slot.name = "netP",
    do.group = FALSE,
    degree.only = TRUE
  )
}

cellchat.control.contact <- compute_multi_single_pathway_centrality(cellchat.control.contact)
cellchat.disease.contact <- compute_multi_single_pathway_centrality(cellchat.disease.contact)
cellchat.control.secreted <- compute_multi_single_pathway_centrality(cellchat.control.secreted)
cellchat.disease.secreted <- compute_multi_single_pathway_centrality(cellchat.disease.secreted)

celltype_cols <- c(
  "eEVT" = "#E64B35",
  "Endothelial" = "#4DBBD5",
  "EVT" = "#00A087"
)

p_control_contact <- SpatialCellChat::spatialDimPlot(
  cellchat.control.contact,
  point.size = 1
) +
  ggtitle("Control") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_disease_contact <- SpatialCellChat::spatialDimPlot(
  cellchat.disease.contact,
  point.size = 1
) +
  ggtitle("Disease") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_control_secreted <- SpatialCellChat::spatialDimPlot(
  cellchat.control.secreted,
  point.size = 1
) +
  ggtitle("Control") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

p_disease_secreted <- SpatialCellChat::spatialDimPlot(
  cellchat.disease.secreted,
  point.size = 1
) +
  ggtitle("Disease") +
  scale_color_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE) +
  scale_fill_manual(values = celltype_cols, limits = names(celltype_cols), drop = FALSE)

plot_multi_pathway_score <- function(p_spatial, chat, pathway.show, point.size = 1) {
  pathway.available <- if ("pathways.cell" %in% names(chat@netP)) {
    chat@netP$pathways.cell
  } else {
    chat@netP$pathways
  }
  
  if (!pathway.show %in% pathway.available) {
    message(pathway.show, " is not present in this CellChat object.")
    return(NULL)
  }
  
  g1 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    measure = c("outdeg"),
    do.binary = FALSE,
    color.heatmap = "Blues",
    point.size = point.size
  )
  
  g2 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    measure = c("indeg"),
    do.binary = FALSE,
    color.heatmap = "Reds",
    point.size = point.size
  )
  
  g3 <- SpatialCellChat::spatialVisual_scoring(
    chat,
    signaling = pathway.show,
    slot.name = "netP",
    do.group = FALSE,
    merge = TRUE,
    do.binary = FALSE,
    color.heatmap = c("Blues", "Reds", "Purples"),
    point.size = point.size
  )
  
  patchwork::wrap_plots(p_spatial, g1, g2, g3, ncol = 4)
}

check_multi_pathways <- function(chat, pathways) {
  pathway.available <- if ("pathways.cell" %in% names(chat@netP)) {
    chat@netP$pathways.cell
  } else {
    chat@netP$pathways
  }
  
  data.frame(
    pathway = pathways,
    present = pathways %in% pathway.available
  )
}
```

## NOTCH multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_contact, cellchat.control.contact, "NOTCH")
plot_multi_pathway_score(p_disease_contact, cellchat.disease.contact, "NOTCH")
```

## NCAM multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_contact, cellchat.control.contact, "NCAM")
plot_multi_pathway_score(p_disease_contact, cellchat.disease.contact, "NCAM")
```

## PD-L1 multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_contact, cellchat.control.contact, "PD-L1")
plot_multi_pathway_score(p_disease_contact, cellchat.disease.contact, "PD-L1")
```

## TNF multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "TNF")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "TNF")
```

## IL1 multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "IL1")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "IL1")
```

## VEGF multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "VEGF")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "VEGF")
```

## LEP multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "LEP")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "LEP")
```

## IL2 multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "IL2")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "IL2")
```

## IL4 multi-sample pathway scores

```{r, fig.width=16, fig.height=4}
plot_multi_pathway_score(p_control_secreted, cellchat.control.secreted, "IL4")
plot_multi_pathway_score(p_disease_secreted, cellchat.disease.secreted, "IL4")
```

